---
title: "Chapter 01. What's in This Book (Read This First!)"
author: "A Solomon Kurz"
date: "`r format(Sys.Date())`"
output:
  github_document
---

# What's in This Book (Read This First!)

## Real people can read this ~~book~~ project based on a book

Kruschke began his text with "This book explains how to actually do Bayesian data analysis, by real people (like you), for realistic data (like yours)." Agreed. Similarly, this project is designed to help those real people do Bayesian data analysis. While I'm at it, I may as well explicate my assumptions about you.

If you're looking at this project, I'm guessing you’re a graduate student, a post-graduate academic or researcher of some sort. Which means I'm presuming you have at least a 101-level foundation in statistics. In his text, it seems like Kruschke presumed his readers would have a good foundation in calculus. I make no such presumption. But if you stats 101 chops are rusty, check out Legler and Roback's free **bookdown** text, [*Broadening Your Statistical Horizons*](https://bookdown.org/roback/bookdown-bysh/). 

I presume a basic working fluency in **R** and a vague idea about what the **tidyverse** is. Kruschke does some **R** warm-up in chapter 2, and I follow suit. But if you're totally new to **R**, you might also consider starting with Peng’s [*R Programming for Data Science*](https://bookdown.org/rdpeng/rprogdatascience/). The best introduction to the **tidyvese**-style of data analysis I've found is Grolemund and Wickham's [*R for Data Science*](http://r4ds.had.co.nz).

## What's in this ~~book~~ project

This project is not meant to stand alone. It's a supplement to the second edition of [Kruschke's *Doing Bayesian Data Analysis*](https://sites.google.com/site/doingbayesiandataanalysis/). I follow the structure of his text, chapter by chapter, translating his analyses into **brms** and **tidyverse** code. However, many of the sections in the text are composed entirely of equations and prose, leaving us nothing to translate. When we run into those sections, the corresponding sections in this project might be blank or even missing.

I reproduce the bulk of the figures in the text, too. However, I don't always reproduce them exactly. I'm passionate about data visualization and like to play around with [color palettes](https://github.com/EmilHvitfeldt/r-color-palettes), formatting templates, and other conventions from time to time. For more on visualization, chapters [3](http://r4ds.had.co.nz/data-visualisation.html), [7](http://r4ds.had.co.nz/exploratory-data-analysis.html), and [28](http://r4ds.had.co.nz/graphics-for-communication.html) in *R4DS* or Healy's [*Data Visualization: A practical introduction*](https://socviz.co).

In this project, I use a handful of formatting conventions gleaned from [*R4DS*](http://r4ds.had.co.nz/introduction.html#running-r-code) and [*R Markdown: The Definitive Guide*](https://bookdown.org/yihui/rmarkdown/software-info.html). 

* I put **R** and **R** packages (e.g., **brms**) in **boldface**.
* **R** code blocks and their output appear in a gray background. E.g., 

```{r}
2 + 2
```

* Did you notice how there were two strips of gray background, there? The first one designated the actual code. The second one was the output of that code. The output of a code block often begins with `##`.
* Functions are in a typewriter font and followed by parentheses, all atop a gray background (e.g., `brm()`).
* When I want to make explicit what packages a given function comes from, I insert the double-colon operator `::` between the package name and the function (e.g., `tidyr::pivot_longer()`).
* **R** objects, such as data or function arguments, are in typewriter font atop a gray background (e.g., `d` or `size = 2`).
* Hyperlinks are denoted by their typical [blue-colored font](https://www.afhayes.com/introduction-to-mediation-moderation-and-conditional-process-analysis.html).

## What's new in the second edition

This is my first attempt at this project. There's nothing new from my end.

## Gimme feedback (be polite)

I am not a statistician and I have no formal background in computer science. In 2018 I just finished my PhD in clinical psychology. During my graduate training I developed an unexpected interest in applied statistics and, more recently, programming. I became an **R** user in 2015 and started learning about Bayesian statistics around 2013. There is still so much to learn, so my apologies for when my code appears dated or inelegant. Thera will also be occasions in which I'm not yet sure how to reproduce models or plots in the text. Which is all to say, [suggestions on how to improve my code are welcome](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/issues).

## Thank you!

While in grad school, I benefitted tremendously from free online content. This project and others like it (e.g., [here](https://bookdown.org/connect/#/apps/1850/access) or [here](https://solomonkurz.netlify.com/post/robust-linear-regression-with-the-robust-student-s-t-distribution/) or [here](https://solomonkurz.netlify.com/post/stein-s-paradox-and-what-partial-pooling-can-do-for-you/)) are my attempts to pay it forward. As soon as you've gained a little proficiency, do consider doing to same. 

I addition to great texts like Kruschke’s, I'd like to point out a few other important resources that have allowed me to complete a project like this:

* Jenny Bryan's [*Happy Git and GitHub for the useR*](http://happygitwithr.com) is the reference that finally got me working on Github.
* Again and again, I return to Grolemund and Wickham's [*R for Data Science*](http://r4ds.had.co.nz) to learn about the **tidyverse** way of coding.
* Yihui Xie's [*bookdown: Authoring Books and Technical Documents with R Markdown*](https://bookdown.org/yihui/bookdown/) is the primary source from which I learned how to make online books like this.

If you haven't already, bookmark these resources and share them with your friends.

## Reference {-}

Kruschke, J. K. (2015). *Doing Bayesian data analysis, Second Edition: A tutorial with R, JAGS, and Stan.* Burlington, MA: Academic Press/Elsevier.

## Session info {-}

At the end of every chapter, I use the `sessionInfo()` function to help make my results more reproducible.

```{r}
sessionInfo()
```

