# Metric Predicted Variable with One Metric Predictor {#sec-Metric-Predicted-Variable-with-One-Metric-Predictor}

> We will initially describe the relationship between the predicted variable, $y$ and predictor, $x$, with a simple linear model and normally distributed residual randomness in $y$. This model is often referred to as 'simple linear regression.' We will generalize the model in three ways. First, we will give it a noise distribution that accommodates outliers, which is to say that we will replace the normal distribution with a $t$ distribution as we did in the previous chapter. The model will be implemented in [**brms**]. Next, we will consider differently shaped relations between the predictor and the predicted, such as quadratic trend. Finally, we will consider hierarchical models of situations in which every individual has data that can be described by an individual trend, and we also want to estimate group-level typical trends across individuals. [@kruschkeDoingBayesianData2015, p. 478]

## Simple linear regression

Load the primary packages for this chapter.

```{r}
#| echo: false
#| eval: false

devtools::install_github("hadley/emo")
```

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(beyonce)
library(patchwork)
library(brms)
library(tidybayes)
library(bayesplot)
```

It wasn't entirely clear how Kruschke simulated the bimodal data on the right panel of Figure 17.1. I figured an even split of two Gaussians would suffice and just sighted their $\mu$'s and $\sigma$'s.

```{r}
# How many draws per panel would you like?
n_draw <- 1000

set.seed(17)
d <- tibble(panel = rep(letters[1:2], each = n_draw),
            x = c(runif(n = n_draw, min = -10, max = 10),
                  rnorm(n = n_draw / 2, mean = -7, sd = 2),
                  rnorm(n = n_draw / 2, mean =  3, sd = 2))) |> 
  mutate(y = 10 + 2 * x + rnorm(n = n_draw, mean = 0, sd = 2))

head(d)
```

In case you missed it, Kruschke defied the formula for these data in Figure 17.1. It is

\begin{align*}
y_i & \sim \operatorname{Normal}(\mu_i, \sigma = 2), \text{where} \\
\mu_i & = 10 + 2 x_i.
\end{align*}

"Note that the model only specifies the dependency of $y$ on $x$. The model does not say anything about what generates $x$, and there is no probability distribution assumed for describing $x$" (p. 479). Let this sink into your soul. It took a long time, for me. E.g., a lot of people fret over the distributions of their $x$ variables. Now one might should examine them to make sure nothing looks off, such as for data coding mistakes. But if they're not perfectly or even approximately Gaussian, that isn't necessarily an issue. The typical linear model makes no presumption about the distribution of the predictors. Often times, the largest issue is whether the $x$ variables are categorical or continuous.

Before we make our Figure 17.1, we'll want to make a separate tibble of the values necessary to plot those sideways Gaussians. Here are the steps.

```{r}
# Define the 3 x-values we want the Gaussians to originate from
curves <- tibble(x = seq(from = -7.5, to = 7.5, length.out = 4)) |>
  # Use the formula 10 + 2x to compute the expected y-value for `x`
  mutate(y_mean = 10 + (2 * x)) |>
  # Based on a Gaussian with `mean = y_mean` and `sd = 2`, compute the 99% intervals
  mutate(ll = qnorm(p = 0.005, mean = y_mean, sd = 2),
         ul = qnorm(p = 0.995, mean = y_mean, sd = 2)) |>
  # Now use those interval bounds to make a sequence of `y` values
  mutate(y = map2(.x = ll, .y = ul, 
                  .f = \(ll, ul) seq(from = ll, to = ul, length.out = 100))) |>
  # Since that operation returned a nested column, we need to `unnest()`
  unnest(y) |>
  # Compute the density values
  mutate(density = map2_dbl(.x = y, .y = y_mean, 
                            .f = \(y, y_mean) dnorm(x = y, mean = y_mean, sd = 2))) |>
  # Now rescale the density values to be wider. Since we want these to 
  # be our `x` values, we'll just redefine the `x` column with these results
  mutate(x = x - density * 2 / max(density))

str(curves)
```

Before we make Figure 17.1, let's talk color. Like last chapter, we'll take our color palette from the **beyonce** package. Our palette will be a nine-point version of #41.

```{r}
#| fig-width: 3
#| fig-height: 1

bp <- beyonce_palette(41, n = 9, type = "continuous")

bp
```

The global theme will be `ggplot2::theme_linedraw()` with the grid lines removed. Make Figure 17.1.

```{r}
#| fig-width: 7
#| fig-height: 4

theme_set(
  theme_linedraw() +
    theme(panel.grid = element_blank())
)

d |> 
  ggplot(aes(x = x, y = y)) +
  geom_vline(xintercept = 0, color = bp[9], linetype = 2, linewidth = 1/3) +
  geom_hline(yintercept = 0, color = bp[9], linetype = 2, linewidth = 1/3) +
  geom_point(alpha = 1/3, color = bp[5], size = 1/3) +
  stat_smooth(method = "lm", formula = 'y ~ x', fullrange = TRUE, se = FALSE, color = bp[1]) +
  geom_path(data = curves,
            aes(group = y_mean),
            color = bp[2], linewidth = 1) +
  labs(title = "Normal PDF around Linear Function",
       subtitle = "We simulated x from a uniform distribution in the left panel and simulated it from a mixture of\n two Gaussians on the right.") +
  coord_cartesian(xlim = c(-10, 10),
                  ylim = c(-10, 30)) +
  theme(strip.background = element_blank(),
        strip.text = element_blank()) +
  facet_wrap(~ panel)
```

Concerning causality,

> *the simple linear model makes no claims about causal connections between* $x$ *and* $y$. *The simple linear model merely describes a tendency for* $y$ *values to be linearly related to* $x$ *values*, hence "predictable" from the $x$ values. When describing data with this model, we are starting with a scatter plot of points generated by an unknown process in the real world, and estimating parameter values that would produce a smattering of points that might mimic the real data. Even if the descriptive model mimics the data well (and it might not), the mathematical "process" in the model may have little if anything to do with the real-world process that created the data. Nevertheless, the parameters in the descriptive model are meaningful because they describe tendencies in the data. (p. 479, *emphasis* added)

I emphasized these points because I've heard and seen a lot of academics conflate linear regression models with causal models. For sure, it might well be preferable if your regression model was also a causal model. But good old prediction is fine, too.

## Robust linear regression {#sec-Robust-linear-regression}

> There is no requirement to use a normal distribution for the noise distribution. The normal distribution is traditional because of its relative simplicity in mathematical derivations. But real data may have outliers, and the use of (optionally) heavy-tailed noise distributions is straight forward in contemporary Bayesian software[, like **brms**]. (pp. 479--480)

Let's make our version of the model diagram in Figure 17.2 to get a sense of where we're going.

```{r}
#| message: false
#| fig-width: 6.75
#| fig-height: 5

# Normal density
p1 <- tibble(x = seq(from = -3, to = 3, by = 0.1)) |> 
  ggplot(aes(x = x, y = (dnorm(x)) / max(dnorm(x)))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 0, y = 0.2,
           label = "normal",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = c(0, 1.5), y = 0.6,
           label = c("italic(M)[0]", "italic(S)[0]"), 
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# A second normal density
p2 <- tibble(x = seq(from = -3, to = 3, by = 0.1)) |> 
  ggplot(aes(x = x, y = (dnorm(x)) / max(dnorm(x)))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 0, y = 0.2,
           label = "normal",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = c(0, 1.5), y = 0.6,
           label = c("italic(M)[1]", "italic(S)[1]"), 
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

## Two annotated arrows
# Save our custom arrow settings
my_arrow <- arrow(angle = 20, length = unit(0.35, "cm"), type = "closed")
p3 <- tibble(x = c(0.33, 1.67),
             y = c(1, 1),
             xend = c(0.75, 1.1),
             yend = c(0, 0)) |>
  ggplot(aes(x = x, xend = xend,
             y = y, yend = yend)) +
  geom_segment(arrow = my_arrow, color = bp[1]) +
  annotate(geom = "text",
           x = c(0.4, 1.25), y = 0.5,
           label = "'~'",
           color = bp[1], family = "Times", parse = TRUE, size = 10) +
  xlim(0, 2) +
  theme_void()

# Exponential density
p4 <- tibble(x = seq(from = 0, to = 1, by = 0.01)) |> 
  ggplot(aes(x = x, y = (dexp(x, 2) / max(dexp(x, 2))))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 0.5, y = 0.2,
           label = "exp",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = 0.5, y = 0.6,
           label = "italic(K)",
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# Likelihood formula
p5 <- tibble(x = 0.5,
             y = 0.25,
             label = "beta[0]+beta[1]*italic(x)[italic(i)]") |> 
  ggplot(aes(x = x, y = y, label = label)) +
  geom_text(color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 1)) +
  ylim(0, 1) +
  theme_void()

# Half-normal density
p6 <- tibble(x = seq(from = 0, to = 3, by = 0.01)) |> 
  ggplot(aes(x = x, y = (dnorm(x)) / max(dnorm(x)))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 1.5, y = 0.2,
           label = "half-normal",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = 1.5, y = 0.6,
           label = "0*','*~italic(S)[sigma]", 
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# Four annotated arrows
p7 <- tibble(x = c(0.43, 0.43, 1.5, 2.5),
             y = c(1, 0.55, 1, 1),
             xend = c(0.43, 1.225, 1.5, 1.75),
             yend = c(0.8, 0.15, 0.2, 0.2)) |>
  ggplot(aes(x = x, xend = xend,
             y = y, yend = yend)) +
  geom_segment(arrow = my_arrow, color = bp[1]) +
  annotate(geom = "text",
           x = c(0.3, 0.7, 1.38, 2), y = c(0.92, 0.22, 0.65, 0.6),
           label = c("'~'", "'='", "'='", "'~'"),
           color = bp[1], family = "Times", parse = TRUE, size = 10) +
  annotate(geom = "text",
           x = 0.43, y = 0.7,
           label = "nu*minute+1",
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  xlim(0, 3) +
  theme_void()

# Student-t density
p8 <- tibble(x = seq(from = -3, to = 3, by = 0.1)) |> 
  ggplot(aes(x = x, y = (dt(x, 3) / max(dt(x, 3))))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 0, y = 0.2,
           label = "student t",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = 0, y = 0.6,
           label = "nu~~~mu[italic(i)]~~~sigma",
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# The final annotated arrow
p9 <- tibble(x = c(0.375, 0.625),
             y = c(1/3, 1/3),
             label = c("'~'", "italic(i)")) |> 
  ggplot(aes(x = x, y = y, label = label)) +
  geom_text(color = bp[1], family = "Times", parse = TRUE, size = c(10, 7)) +
  geom_segment(x = 0.5, xend = 0.5,
               y = 1, yend = 0, 
               arrow = my_arrow, color = bp[1]) +
  xlim(0, 1) +
  theme_void()

# Some text
p10 <- tibble(x = 0.5,
              y = 0.5,
              label = "italic(y[i])") |> 
  ggplot(aes(x = x, y = y, label = label)) +
  geom_text(color = bp[1], family = "Times", parse = TRUE, size = 7) +
  xlim(0, 1) +
  theme_void()

# Define the layout
layout <- c(
  area(t = 1, b = 2, l = 3, r = 5),
  area(t = 1, b = 2, l = 7, r = 9),
  area(t = 4, b = 5, l = 1, r = 3),
  area(t = 4, b = 5, l = 5, r = 7),
  area(t = 4, b = 5, l = 9, r = 11),
  area(t = 3, b = 4, l = 3, r = 9),
  area(t = 7, b = 8, l = 5, r = 7),
  area(t = 6, b = 7, l = 1, r = 11),
  area(t = 9, b = 9, l = 5, r = 7),
  area(t = 10, b = 10, l = 5, r = 7))

# Combine and plot!
(p1 + p2 + p4 + p5 + p6 + p3 + p8 + p7 + p9 + p10) + 
  plot_layout(design = layout) &
  ylim(0, 1) &
  theme(plot.margin = margin(0, 5.5, 0, 5.5))
```

As introduced in @sec-The-steps-of-Bayesian-data-analysis, here's the code for the `height_weight_generator()` function.

```{r}
height_weight_generator <- function(n, male_prob = 0.5, seed = 47405) {

  # Male MVN distribution parameters
  height_male_mu <- 69.18;    height_male_sd <- 2.87
  log_weight_male_mu <- 5.14; log_weight_male_sd <- 0.17
  male_rho <- 0.42
  male_mean <- c(height_male_mu, log_weight_male_mu)
  male_sigma <- matrix(c(height_male_sd^2, male_rho * height_male_sd * log_weight_male_sd,
                         male_rho * height_male_sd * log_weight_male_sd, log_weight_male_sd^2),
                       nrow = 2)

  # Female cluster 1 MVN distribution parameters
  height_female_mu_1 <- 63.11;    height_female_sd_1 <- 2.76
  log_weight_female_mu_1 <- 5.06; log_weight_female_sd_1 <- 0.24
  female_rho_1 <- 0.41
  female_mean_1 <- c(height_female_mu_1, log_weight_female_mu_1)
  female_sigma_1 <- matrix(c(height_female_sd_1^2, female_rho_1 * height_female_sd_1 * log_weight_female_sd_1,
                             female_rho_1 * height_female_sd_1 * log_weight_female_sd_1, log_weight_female_sd_1^2),
                           nrow = 2)

  # Female cluster 2 MVN distribution parameters
  height_female_mu_2 <- 64.36;    height_female_sd_2 <- 2.49
  log_weight_female_mu_2 <- 4.86; log_weight_female_sd_2 <- 0.14
  female_rho_2 <- 0.44
  female_mean_2 <- c(height_female_mu_2, log_weight_female_mu_2)
  female_sigma_2 <- matrix(c(height_female_sd_2^2, female_rho_2 * height_female_sd_2 * log_weight_female_sd_2,
                             female_rho_2 * height_female_sd_2 * log_weight_female_sd_2, log_weight_female_sd_2^2),
                           nrow = 2)

  # Female cluster probabilities
  female_prob_1 <- 0.46
  female_prob_2 <- 1 - female_prob_1

  # Randomly generate data values from those MVN distributions
  if (!is.null(seed)) set.seed(seed)
  datamatrix <- matrix(0, nrow = n, ncol = 3)
  colnames(datamatrix) <- c("male", "height", "weight")
  male_val <- 1; female_val <- 0  # Arbitrary coding values
  for (i in 1:n) {
    # Generate `sex`
    sex <- sample(c(male_val, female_val), size = 1, replace = TRUE, prob = c(male_prob, 1 - male_prob))
    if (sex == male_val) datum <- MASS::mvrnorm(n = 1, mu = male_mean, Sigma = male_sigma)
    if (sex == female_val) {
      female_cluster <- sample(c(1, 2), size = 1, replace = TRUE, prob = c(female_prob_1, female_prob_2))
      if (female_cluster == 1) datum <- MASS::mvrnorm(n = 1, mu = female_mean_1, Sigma = female_sigma_1)
      if (female_cluster == 2) datum <- MASS::mvrnorm(n = 1, mu = female_mean_2, Sigma = female_sigma_2)
    }
    datamatrix[i, ] <- c(sex, round(c(datum[1], exp(datum[2])), digits = 1))
  }

  data.frame(datamatrix)
}
```

Let's take this baby for a spin to simulate our data.

```{r}
d <- height_weight_generator(n = 300, male_prob = 0.50, seed = 17) |> 
  # This will allow us to subset 30 of the values into their own group
  mutate(subset = rep(0:1, times = c(9, 1)) |> rep(times = 30))

head(d)
```

Note how we set the seed for the the pseudorandom number generator with the `seed` argument, which makes the results in this ebook reproducible. But since we do not know what seed value  Kruschke used to simulate the data in this section of the test, the results of our models in the next section will differ a little from those in the text. However, if you want to more closely reproduce Kruschke's examples, load the `HtWtData30.csv` and `HtWtData300.csv` data files and fit the models to those, instead.

```{r}
#| echo: false
#| eval: false

Data30 <- read_csv("data.R/HtWtData30.csv")
Data300 <- read_csv("data.R/HtWtData300.csv")
```

Anyway,

> fortunately, we do not have to worry much about analytical derivations because we can let JAGS or Stan generate a high resolution picture of the posterior distribution. Our job, therefore, is to specify sensible priors and to make sure that the MCMC process generates a trustworthy posterior sample that is converged and well mixed. (p. 483)

### Robust linear regression in ~~JAGS~~ brms

Presuming a data set with a sole standardized predictor `x_z` for a sole standardized criterion `y_z`, the basic **brms** code corresponding to the JAGS code Kruschke showed on page 483 looks like this.

```{r}
#| eval: false

fit <- brm(
  data = my_data,
  family = student,
  y_z ~ 1 + x_z,
  prior = c(prior(normal(0, 10), class = Intercept),
            prior(normal(0, 10), class = b),
            prior(normal(0, 1), class = sigma),
            prior(exponential(one_over_twentynine), class = nu)),
  stanvars = stanvar(1/29, name = "one_over_twentynine"))
```

Like we discussed in Chapter 16, we won't be using the uniform prior for $\sigma$. Since we're presuming standardized data, a half-unit normal is a fine choice. But do note this is much tighter than Kruschke's $\operatorname{Uniform}(0.001, 1000)$ and it will have down-the-road consequences for our results versus those in the text.

Also, look at how we just pumped the definition of our sole `stanvar(1/29, name = "one_over_twentynine")` operation right into the `stanvar` argument. If we were defining multiple values this way, I'd prefer to save this as an object first and then just pump that object into `stanvars`. But in this case, it was simple enough to just throw directly into the `brm()` function.

#### Standardizing the data for MCMC sampling

"Standardizing simply means re-scaling the data relative to their mean and standard deviation" (p. 485). For any variable $x$, that follows the formula

$$z_i = \frac{x_i - \bar x}{s},$$

where $x_i$ is the $i$th row in the vector of $x$ values, $\bar x$ is the sample mean for $x$, $s$ is the sample standard deviation for $x$, and the product of the standardizing procedure is $z_i$ (i.e., the $z$-score).

Kruschke mentioned how standardizing your data before feeding it into JAGS often helps the algorithm operate smoothly. The same basic principle holds for **brms** and Stan. Stan can often handle unstandardized data just fine. But if you ever run into estimation difficulties, consider standardizing your data and trying again.

We'll make a simple function to standardize the `height` and `weight` values.

```{r}
standardize <- function(x) {
  (x - mean(x)) / sd(x)
}

d <- d |> 
  mutate(height_z = standardize(height),
         weight_z = standardize(weight))
```

Somewhat analogous to how Kruschke standardized his data within the JAGS code, you could standardize the data within the `brm()` function. That would look something like this.

```{r}
#| eval: false

fit <- brm(
  data = d |>  # The standardizing occurs in the next two lines
    mutate(height_z = standardize(height),
           weight_z = standardize(weight)),
  family = student,
  weight_z ~ 1 + height_z)
```

I code like this once in a while, but I generally don't recommend it. If you feel the need to transform the variables in your model, it might be safer to save those transformed versions in a data frame that is accessible outside of the model.

We'll fit the two models at once. `fit1` will be of the total data sample. `fit2` is of the $n = 30$ subset.

```{r fit17.1}
fit17.1 <- brm(
  data = d,
  family = student,
  weight_z ~ 1 + height_z,
  prior = c(prior(normal(0, 10), class = Intercept),
            prior(normal(0, 10), class = b),
            prior(normal(0, 1), class = sigma),
            prior(exponential(one_over_twentynine), class = nu)),
  chains = 4, cores = 4,
  stanvars = stanvar(1/29, name = "one_over_twentynine"),
  seed = 17,
  file = "fits/fit17.01")

fit17.2 <- update(
  fit17.1,
  newdata = d |> 
    filter(subset == 1),
  chains = 4, cores = 4,
  seed = 17,
  file = "fits/fit17.02")
```

```{r}
#| echo: false
#| eval: false

Data30 <- read_csv("data.R/HtWtData30.csv")
Data300 <- read_csv("data.R/HtWtData300.csv")

Data30 <- Data30 |> 
  mutate(height_z = standardize(height),
         weight_z = standardize(weight))

Data300 <- Data300 |> 
  mutate(height_z = standardize(height),
         weight_z = standardize(weight))

fit17.1b <- brm(
  data = Data300,
  family = student,
  weight_z ~ 1 + height_z,
  prior = c(prior(normal(0, 10), class = Intercept),
            prior(normal(0, 10), class = b),
            prior(normal(0, 1), class = sigma),
            prior(exponential(one_over_twentynine), class = nu)),
  chains = 4, cores = 4,
  stanvars = stanvar(1/29, name = "one_over_twentynine"),
  seed = 17)

fit17.2b <- update(
  fit17.1b,
  newdata = Data30,
  chains = 4, cores = 4,
  seed = 17)

print(fit17.1)
print(fit17.1b)

print(fit17.2)
print(fit17.2b)
```

Here are the results.

```{r}
print(fit17.1)
print(fit17.2)
```

Based on Kruschke's Equation 17.2, we can convert the standardized coefficients back to their original metric using the formulas

\begin{align*}
\beta_0 & = \zeta_0 \operatorname{SD}_y + M_y - \frac{\zeta_1 M_x \operatorname{SD}_y}{\operatorname{SD}_x} \;\; \text{and}  \\
\beta_1 & = \frac{\zeta_1 \operatorname{SD}_y}{\operatorname{SD}_x},
\end{align*}

where $\zeta_0$ and $\zeta_1$ denote the intercept and slope for the model of the standardized data, respectively, and that model follows the familiar form

$$z_{\hat y} = \zeta_0 + \zeta_1 z_x.$$

To implement those equations, we'll first extract the posterior draws. We begin with `fit17.1`, the model for which $N = 300$.

```{r}
draws <- as_draws_df(fit17.1)

head(draws)
```

Let's wrap the consequences of Equation 17.2 into two functions.

```{r}
make_beta_0 <- function(zeta_0, zeta_1, sd_x, sd_y, m_x, m_y) {
  zeta_0 * sd_y + m_y - zeta_1 * m_x * sd_y / sd_x
}

make_beta_1 <- function(zeta_1, sd_x, sd_y) {
  zeta_1 * sd_y / sd_x
}
```

After saving a few values, we're ready to use our custom functions to convert our posteriors for `b_Intercept` and `b_height_z` to their natural metric.

```{r}
# Save sample statistics
sd_x <- sd(d$height)
sd_y <- sd(d$weight)
m_x  <- mean(d$height)
m_y  <- mean(d$weight)

# Wrangle
draws <- draws |> 
  mutate(b_0 = make_beta_0(zeta_0 = b_Intercept,
                           zeta_1 = b_height_z,
                           sd_x   = sd_x,
                           sd_y   = sd_y,
                           m_x    = m_x,
                           m_y    = m_y),
         b_1 = make_beta_1(zeta_1 = b_height_z,
                           sd_x   = sd_x,
                           sd_y   = sd_y))

glimpse(draws)
```

Now we're finally ready to make the top panel of Figure 17.4.

```{r}
#| fig-width: 4.5
#| fig-height: 4

# How many posterior lines would you like?
n_lines <- 100

d |> 
  ggplot(aes(x = height, y = weight)) +
  geom_abline(data = draws |> slice(1:n_lines),
              aes(intercept = b_0, slope = b_1, group = .draw),
              alpha = 1/3, color = bp[2], linewidth = 1/4) + 
  geom_point(alpha = 1/2, color = bp[5]) +
  labs(x = "height",
       y = "weight",
       subtitle = eval(substitute(paste("Data with", n_lines, "credible regression lines")))) +
  coord_cartesian(xlim = c(50, 80),
                  ylim = c(-50, 470))
```

Now we make the histograms.

```{r}
#| warning: false
#| fig-width: 6
#| fig-height: 4

# We'll use this to mark off the ROPEs as white strips in the background
rope <- tibble(name = "Slope", 
               xmin = -0.5,
               xmax = 0.5)

# Annotate the ROPE
d_text <- tibble(x = 0,
                 y = 0.98,
                 label = "ROPE",
                 name  = "Slope")

# Here are the primary data
draws |> 
  transmute(Intercept = b_0,
            Slope     = b_1,
            Scale     = sigma * sd_y,
            Normality = nu |> log10()) |> 
  pivot_longer(cols = everything()) |> 
  
  # Plot
  ggplot() +
  geom_rect(data = rope,
            aes(xmin = xmin, xmax = xmax,
                ymin = -Inf, ymax = Inf),
            color = "transparent", fill = bp[9]) +
  stat_histinterval(aes(x = value), 
                    point_interval = mode_hdi, .width = 0.95,
                    breaks = 40, color = bp[1], fill = bp[6],
                    normalize = "panels", slab_color = bp[5]) +
  geom_text(data = d_text,
            aes(x = x, y = y, label = label),
            color = "white", size = 2.75) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(NULL) +
  facet_wrap(~ name, scales = "free", ncol = 2)
```

Note how we used the formula

$$\sigma_y = \sigma_z \operatorname{SD}_y$$

to transform the residual variance from the standardized model ($\sigma_z$) to a metric on the original scale of the data ($\sigma_y$). You can get that formula at the bottom of page 485 in the text. Given its simplicity, we didn't bother wrapping that formula into a function, but you could if you wanted to.

Here's the scatter plot for the slope and intercept.

```{r}
#| fig-width: 3.25
#| fig-height: 3

draws |> 
  ggplot(aes(x = b_1, y = b_0)) +
  geom_point(alpha = 1/3, color = bp[3], size = 1/3) +
  labs(x = expression(beta[1]),
       y = expression(beta[0]))
```

That is one strong correlation! Finally, here's the scatter plot for $\operatorname{log_{10}}(\nu)$ and $\sigma_{\text{raw metric}}$.

```{r}
#| warning: false
#| fig-width: 3.25
#| fig-height: 3

draws |> 
  transmute(Scale     = sigma * sd_y,
            Normality = nu |> log10()) |> 
  ggplot(aes(x = Normality, y = Scale)) +
  geom_point(color = bp[3], size = 1/3, alpha = 1/3) +
  labs(x = expression(log[10](nu)),
       y = expression(sigma))
```

Let's back track and make the plots for Figure 17.3 with `fit17.2`. We'll need to extract the posterior draws and wrangle, as before.

```{r}
draws <- as_draws_df(fit17.2)

draws <- draws |> 
  mutate(b_0 = make_beta_0(zeta_0 = b_Intercept,
                           zeta_1 = b_height_z,
                           sd_x   = sd_x,
                           sd_y   = sd_y,
                           m_x    = m_x,
                           m_y    = m_y),
         b_1 = make_beta_1(zeta_1 = b_height_z,
                           sd_x   = sd_x,
                           sd_y   = sd_y))

glimpse(draws)
```

Here's the top panel of Figure 17.3.

```{r}
#| fig-width: 4.5
#| fig-height: 4

# How many posterior lines would you like?
n_lines <- 100

ggplot(data = d |> 
         filter(subset == 1), 
       aes(x = height, y = weight)) +
  geom_vline(xintercept = 0, color = bp[9]) +
  geom_abline(data = draws |> slice(1:n_lines),
              aes(intercept = b_0, slope = b_1, group = .draw),
              alpha = 1/3, color = bp[6], linewidth = 1/4) +
  geom_point(alpha = 1/2, color = bp[3]) +
  scale_y_continuous(breaks = seq(from = -300, to = 200, by = 100)) +
  labs(x = "height",
       y = "weight",
       subtitle = eval(substitute(paste("Data with", n_lines, "credible regression lines")))) +
  coord_cartesian(xlim = c(0, 80),
                  ylim = c(-350, 250))
```

Next we'll make the histograms.

```{r}
#| warning: false
#| fig-width: 6
#| fig-height: 4

# Here are the primary data
draws |> 
  transmute(Intercept = b_0,
            Slope     = b_1,
            Scale     = sigma * sd_y,
            Normality = nu |> log10()) |> 
  pivot_longer(cols = everything()) |> 
  
  # Plot
  ggplot() +
  geom_rect(data = rope,
            aes(xmin = xmin, xmax = xmax,
                ymin = -Inf, ymax = Inf),
            color = "transparent", fill = bp[9]) +
  stat_histinterval(aes(x = value), 
                    point_interval = mode_hdi, .width = 0.95,
                    breaks = 40, color = bp[1], fill = bp[6],
                    normalize = "panels", slab_color = bp[5]) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(NULL) +
  facet_wrap(~ name, scales = "free", ncol = 2)
```

And we'll finish up with the scatter plots.

```{r}
#| warning: false
#| fig-width: 3.25
#| fig-height: 3

draws |> 
  ggplot(aes(x = b_1, y = b_0)) +
  geom_point(alpha = 1/3, color = bp[4], size = 1/3) +
  labs(x = expression(beta[1]),
       y = expression(beta[0]))

draws |> 
  transmute(Scale = sigma * sd_y,
            Normality = nu |> log10()) |> 
  ggplot(aes(x = Normality, y = Scale)) +
  geom_point(alpha = 1/3, color = bp[4], size = 1/3) +
  labs(x = expression(log[10](nu)),
       y = expression(sigma))
```

### Robust linear regression in Stan

> Recall from @sec-HMC-sampling (p. 400) that Stan uses Hamiltonian dynamics to find proposed positions in parameter space. The trajectories use the gradient of the posterior distribution to move large distances even in narrow distributions. Thus, HMC by itself, without data standardization, should be able to efficiently generate a representative sample from the posterior distribution. (p. 487)

To be clear, we're going to fit the models with Stan/**brms** twice. Above, we used the standardized data like Kruschke did with his JAGS code. Now we're getting ready to follow along with the text and use Stan/**brms** to fit the models with the *un*standardized data.

#### Constants for vague priors

The issues is we want a system where we can readily specify vague priors on our regression models when the data are not standardized. As it turns out,

> a regression slope can take on a maximum value of $\operatorname{SD}_y / \operatorname{SD}_x$ for data that are perfectly correlated. Therefore, the prior on the slope will be given a standard deviation that is large compared to that maximum. The biggest that an intercept could be, for data that are perfectly correlated, is $M_x \operatorname{SD}_y / \operatorname{SD}_x$. Therefore, the prior on the intercept will have a standard deviation that is large compared to that maximum. (p. 487)

With that in mind, we'll specify our `stanvars` as follows.

```{r}
beta_0_sigma <- 10 * abs(m_x * sd_y / sd_x)
beta_1_sigma <- 10 * abs(sd_y / sd_x) 

stanvars <- stanvar(beta_0_sigma, name = "beta_0_sigma") + 
  stanvar(beta_1_sigma, name = "beta_1_sigma") +
  stanvar(sd_y, name = "sd_y") +
  stanvar(1/29, name = "one_over_twentynine")
```

As in Chapter 16, "set the priors to be extremely broad relative to the data" (p. 487). With our `stanvars` defined, we're ready to fit `fit17.3`.

```{r fit17.3}
fit17.3 <- brm(
  data = d,
  family = student,
  weight ~ 1 + height,
  prior = c(prior(normal(0, beta_0_sigma), class = Intercept),
            prior(normal(0, beta_1_sigma), class = b),
            prior(normal(0, sd_y), class = sigma),
            prior(exponential(one_over_twentynine), class = nu)),
  chains = 4, cores = 4,
  stanvars = stanvars,
  seed = 17,
  file = "fits/fit17.03")
```

Here's the model summary.

```{r}
print(fit17.3)
```

Now compare the histograms for these posterior draws to those we made, above, from those `fit17.1`. You'll see they're quite similar.

```{r}
#| warning: false
#| fig-width: 6
#| fig-height: 4

# Here are the primary data
as_draws_df(fit17.3) |> 
  transmute(Intercept = b_Intercept,
            Slope     = b_height,
            Scale     = sigma,
            Normality = nu |> log10()) |> 
  pivot_longer(cols = everything()) |> 
  
  # Plot
  ggplot() +
  geom_rect(data = rope,
            aes(xmin = xmin, xmax = xmax,
                ymin = -Inf, ymax = Inf),
            color = "transparent", fill = bp[9]) +
  stat_histinterval(aes(x = value), 
                    point_interval = mode_hdi, .width = 0.95,
                    breaks = 40, color = bp[1], fill = bp[6],
                    normalize = "panels", slab_color = bp[5]) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(NULL) +
  facet_wrap(~ name, ncol = 2, scales = "free")
```

### Stan or JAGS?

In this ebook we only fit the models with **brms**, which uses Stan under the hood. But since we fit the $N = 300$ model with both standardized and unstandardized data, we can compare their performance. They had equally impressive autocorrelation plots.

```{r}
#| fig-width: 5
#| fig-height: 3

# Set the bayesplot color scheme
color_scheme_set(scheme = bp[c(1, 3, 8, 7, 5, 5)])

as_draws_df(fit17.1) |> 
  mutate(chain = .chain) |> 
  mcmc_acf(pars = vars(b_Intercept:nu),
           lags = 10) +
  ggtitle("fit17.1")

as_draws_df(fit17.3) |> 
  mutate(chain = .chain) |> 
  mcmc_acf(pars = vars(b_Intercept:nu),
           lags = 10) +
  ggtitle("fit17.3")
```

Their $N_\textit{eff}/N$ ratios were pretty similar. Both were reasonable. You'd probably want to run a simulation to contrast them with any rigor.

```{r}
#| message: false
#| warning: false
#| fig-width: 6
#| fig-height: 3.25

# Change the bayesplot color scheme
color_scheme_set(scheme = bp[c(1, 3, 4, 6, 7, 9)])

p1 <- neff_ratio(fit17.1) |> 
  mcmc_neff() +
  yaxis_text(hjust = 0) +
  ggtitle("fit17.1")

p2 <- neff_ratio(fit17.3) |> 
  mcmc_neff() +
  yaxis_text(hjust = 0) +
  ggtitle("fit17.3")

p1 / p2 + plot_layout(guide = "collect")
```

### Interpreting the posterior distribution {#sec-Interpreting-the-posterior-distribution}

Halfway through the prose, Kruschke mentioned how the models provide entire posteriors for the `weight` of a 50-inch-tall person. **brms** offers a few ways to do so.

> In some applications, there is interest in extrapolating or interpolating trends at $x$ values sparsely represented in the current data. For instance, we might want to predict the weight of a person who is $50$ inches tall. A feature of Bayesian analysis is that we get an entire distribution of credible predicted values, not only a point estimate. (p. 489)

Since this is such a simple model, one way is to work directly with the posterior draws Here we use the model formula $y_i = \beta_0 + \beta_1 x_i$ by adding the transformed intercept `b_0` to the product of 50 and the transformed coefficient for `height`, `b_1`.

```{r}
#| fig-width: 3.5
#| fig-height: 2.5

draws |> 
  mutate(weight_at_50 = b_0 + b_1 * 50) |> 
  
  ggplot(aes(x = weight_at_50)) +
  stat_histinterval(point_interval = mode_hdi, .width = 0.95,
                    breaks = 40, color = bp[1], fill = bp[6],
                    normalize = "panels", slab_color = bp[5]) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab("lbs")
```

Looks pretty wide, doesn't it? Hopefully this isn't a surprise. Recall that this `draws` is from `fit17.2`, the posterior based on the $n = 30$ data. With so few cases, most predictions from that model are uncertain. But also, 50 inches is way out of the bounds of the data the model was based on, so we should be uncertain in this range.

Let's practice a second method. With the `brms::fitted()` function, we can specify the desired `height` value into a tibble, which we'll then feed into the `newdata` argument. Fitted will then return the model-implied criterion value for that predictor variable. To warm up, we'll first to it with `fit17.3`, the model based on the untransformed data.

```{r}
nd <- tibble(height = 50)

fitted(fit17.3, newdata = nd)
```

The code returned a typical **brms**-style summary of the posterior mean, standard deviation, and 95% percentile-based intervals. The same basic method will work for the standardized models, `fit17.1` or `fit17.2`. But that will take a little more wrangling. First, we'll need to transform our desired value 50 into its standardized version.

```{r}
nd <- tibble(height_z = (50 - mean(d$height)) / sd(d$height))
```

When we feed this value into `fitted()`, it will return the corresponding posterior within the standardized metric. But we want unstandardized, so we'll need to transform. That'll be a few-step process. First, to do the transformation properly we'll want to work with the poster draws themselves, rather than summary values. So we'll set `summary = FALSE`. We'll then convert the draws into a tibble format. Then we'll use the `transmute()` function to do the conversion. In the final step, we'll use `mean_qi()` to compute the summary values.

```{r}
#| message: false
#| warning: false

fitted(fit17.1,
       newdata = nd,
       summary = FALSE) |> 
  as_tibble() |> 
  transmute(weight = V1 * sd(d$weight) + mean(d$weight)) |> 
  mean_qi()
```

If you look above, you'll see the results are within rounding error of those from `fit3`.

## Hierarchical regression on individuals within groups {#sec-Hierarchical-regression-on-individuals-within-groups}

> In the previous applications, the $j$th individual contributed a single $x_j, y_j$ pair. But suppose instead that every individual, $j$, contributes multiple observations of $x_{i \mid j}, y_{i \mid j}$ pairs. (The subscript notation $i \mid j$ means the $i$th observation within the $j$th individual.) With these data, we can estimate a regression curve for every individual. If we also assume that the individuals are mutually representative of a common group, then we can estimate group-level parameters too. (p. 490)

Load the fictitious data and take a `glimpse()`.

```{r}
#| message: false

my_data <- read_csv("data.R/HierLinRegressData.csv")

glimpse(my_data)
```

> Our goal is to describe each individual with a linear regression, and simultaneously to estimate the typical slope and intercept of the group overall. A key assumption for our analysis is that each individual is representative of the group. Therefore, every individual informs the estimate of the group slope and intercept, which in turn inform the estimates of all the individual slopes and intercepts. Thereby we get sharing of information across individuals, and shrinkage of individual estimates toward the overarching mode. (p. 491)

### The model and implementation in ~~JAGS~~ brms {#sec-The-model-and-implementation-in-brms}

Kruschke described the model diagram in Figure 17.6 as "a bit daunting" (p. 491). The code to make our version of the diagram is "a bit daunting," too. Just like the code for any other diagram, it's modular. If you're following along with me and making these on your own, just build it up, step by step.

```{r}
#| message: false
#| fig-width: 8.5
#| fig-height: 6

# Normal density
p1 <- tibble(x = seq(from = -3, to = 3, by = 0.1)) |> 
  ggplot(aes(x = x, y = (dnorm(x)) / max(dnorm(x)))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 0, y = 0.2,
           label = "normal",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = c(0, 1.5), y = 0.6,
           label = c("italic(M)[0]", "italic(S)[0]"), 
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# Half-normal density
p2 <- tibble(x = seq(from = 0, to = 3, by = 0.01)) |> 
  ggplot(aes(x = x, y = (dnorm(x)) / max(dnorm(x)))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 1.5, y = 0.2,
           label = "half-normal",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = 1.5, y = 0.6,
           label = "0*','*~italic(S)[sigma][0]", 
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# A second normal density
p3 <- tibble(x = seq(from = -3, to = 3, by = 0.1)) |> 
  ggplot(aes(x = x, y = (dnorm(x)) / max(dnorm(x)))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 0, y = 0.2,
           label = "normal",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = c(0, 1.5), y = 0.6,
           label = c("italic(M)[1]", "italic(S)[1]"), 
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# A second half-normal density
p4 <- tibble(x = seq(from = 0, to = 3, by = 0.01)) |> 
  ggplot(aes(x = x, y = (dnorm(x)) / max(dnorm(x)))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 1.5, y = 0.2,
           label = "half-normal",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = 1.5, y = 0.6,
           label = "0*','*~italic(S)[sigma][1]", 
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# Four annotated arrows
p5 <- tibble(x = c(0.05, 0.35, 0.65, 0.95),
             y = 1,
             xend = c(0.32, 0.4, 0.65, 0.72),
             yend = 0.2) |>
  ggplot(aes(x = x, xend = xend,
             y = y, yend = yend)) +
  geom_segment(arrow = my_arrow, color = bp[1]) +
  annotate(geom = "text",
           x = c(0.15, 0.35, 0.625, 0.78), y = 0.55,
           label = "'~'",
           color = bp[1], family = "Times", parse = TRUE, size = 10) +
  xlim(0, 1) +
  theme_void()

# Third normal density
p6 <- tibble(x = seq(from = -3, to = 3, by = 0.1)) |> 
  ggplot(aes(x = x, y = (dnorm(x)) / max(dnorm(x)))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 0, y = 0.2,
           label = "normal",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = c(0, 1.5), y = 0.6,
           label = c("mu[0]", "sigma[0]"), 
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# Fourth normal density
p7 <- tibble(x = seq(from = -3, to = 3, by = 0.1)) |> 
  ggplot(aes(x = x, y = (dnorm(x)) / max(dnorm(x)))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 0, y = 0.2,
           label = "normal",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = c(0, 1.5), y = 0.6,
           label = c("mu[1]", "sigma[1]"), 
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# Two annotated arrows
p8 <- tibble(x = c(0.18, 0.82),
             y = 1,
             xend = c(0.36, 0.55),
             yend = 0) |>
  ggplot(aes(x = x, xend = xend,
             y = y, yend = yend)) +
  geom_segment(arrow = my_arrow, color = bp[1]) +
  annotate(geom = "text",
           x = c(0.18, 0.33, 0.64, 0.77), y = 0.55,
           label = c("'~'", "italic(j)", "'~'", "italic(j)"),
           size = c(10, 7, 10, 7), 
           color = bp[1], family = "Times", parse = TRUE) +
  xlim(0, 1) +
  theme_void()

# Exponential density
p9 <- tibble(x = seq(from = 0, to = 1, by = 0.01)) |> 
  ggplot(aes(x = x, y = (dexp(x, 2) / max(dexp(x, 2))))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 0.5, y = 0.2,
           label = "exp",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = 0.5, y = 0.6,
           label = "italic(K)",
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# Likelihood formula
p10 <- tibble(x = 0.5,
              y = 0.25,
              label = "beta[0][italic(j)]+beta[1][italic(j)]*italic(x)[italic(i)*'|'*italic(j)]") |> 
  ggplot(aes(x = x, y = y, label = label)) +
  geom_text(color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 1)) +
  ylim(0, 1) +
  theme_void()

# Half-normal density
p11 <- tibble(x = seq(from = 0, to = 3, by = 0.01)) |> 
  ggplot(aes(x = x, y = (dnorm(x)) / max(dnorm(x)))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 1.5, y = 0.2,
           label = "half-normal",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = 1.5, y = 0.6,
           label = "0*','*~italic(S)[sigma]", 
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# Four annotated arrows
p12 <- tibble(x = c(0.43, 0.43, 1.5, 2.5),
              y = c(1, 0.55, 1, 1),
              xend = c(0.43, 1.225, 1.5, 1.75),
              yend = c(0.8, 0.15, 0.2, 0.2)) |>
  ggplot(aes(x = x, xend = xend,
             y = y, yend = yend)) +
  geom_segment(arrow = my_arrow, color = bp[1]) +
  annotate(geom = "text",
           x = c(0.3, 0.7, 1.38, 2), y = c(0.92, 0.22, 0.65, 0.6),
           label = c("'~'", "'='", "'='", "'~'"),
           size = 10, 
           color = bp[1], family = "Times", parse = TRUE) +
  annotate(geom = "text",
           x = 0.43, y = 0.7,
           label = "nu*minute+1",
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  xlim(0, 3) +
  theme_void()

# Student-t density
p13 <- tibble(x = seq(from = -3, to = 3, by = 0.1)) |> 
  ggplot(aes(x = x, y = (dt(x, 3) / max(dt(x, 3))))) +
  geom_area(fill = bp[6]) +
  annotate(geom = "text",
           x = 0, y = 0.2,
           label = "student t",
           color = bp[1], size = 7) +
  annotate(geom = "text",
           x = 0, y = 0.6,
           label = "nu~~mu[italic(i)*'|'*italic(j)]~~sigma",
           color = bp[1], family = "Times", parse = TRUE, size = 7) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.line.x = element_line(color = bp[1], linewidth = 0.5))

# The final annotated arrow
p14 <- tibble(x = c(0.375, 0.625),
              y = c(1/3, 1/3),
              label = c("'~'", "italic(i)*'|'*italic(j)")) |> 
  ggplot(aes(x = x, y = y, label = label)) +
  geom_text(color = bp[1], family = "Times", parse = TRUE, size = c(10, 7), ) +
  geom_segment(x = 0.5, xend = 0.5,
               y = 1, yend = 0, 
               arrow = my_arrow, color = bp[1]) +
  xlim(0, 1) +
  theme_void()

# Some text
p15 <- tibble(x = 0.5,
              y = 0.5,
              label = "italic(y)[italic(i)*'|'*italic(j)]") |> 
  ggplot(aes(x = x, y = y, label = label)) +
  geom_text(color = bp[1], family = "Times", parse = TRUE, size = 7) +
  xlim(0, 1) +
  theme_void()

# Define the layout
layout <- c(
  area(t = 1, b = 2, l = 1, r = 3),
  area(t = 1, b = 2, l = 5, r = 7),
  area(t = 1, b = 2, l = 9, r = 11),
  area(t = 1, b = 2, l = 13, r = 15),
  area(t = 4, b = 5, l = 5, r = 7),
  area(t = 4, b = 5, l = 9, r = 11),
  area(t = 3, b = 4, l = 1, r = 15),
  area(t = 7, b = 8, l = 3, r = 5),
  area(t = 7, b = 8, l = 7, r = 9),
  area(t = 7, b = 8, l = 11, r = 13),
  area(t = 6, b = 7, l = 5, r = 11),
  area(t = 10, b = 11, l = 7, r = 9),
  area(t = 9, b = 10, l = 3, r = 13),
  area(t = 12, b = 12, l = 7, r = 9),
  area(t = 13, b = 13, l = 7, r = 9))

# Combine, adjust, and display!
(p1 + p2 + p3 + p4 + p6 + p7 + p5 + p9 + p10 + p11 + p8 + p13 + p12 + p14 + p15) + 
  plot_layout(design = layout) &
  ylim(0, 1) &
  theme(plot.margin = margin(0, 5.5, 0, 5.5))
```

Just look at that sweet thing! If you made your own version; have a piece of cake. `r emo::ji("cake")` You earned it.

Now let's standardize the data and define our `stanvars`. I should note that standardizing and mean centering, more generally, becomes complicated with multilevel models. Here we're just standardizing based on the grand mean and grand standard deviation. But there are other ways to standardize, such as within groups. Craig Enders has a good [-@endersCenteringPredictorsContextual2013] [book chapter](https://scholar.google.com/scholar?hl=en&as_sdt=0%2C1&q=Centering+Predictors+and+Contextual+Effects&btnG=) that touched on the topic, as well as an earlier [-@enders2007centering] [paper](http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.928.9848&rep=rep1&type=pdf) with Tofighi.

```{r}
my_data <- my_data |> 
  mutate(x_z = standardize(X),
         y_z = standardize(Y))
```

In my experience, you typically use the `(|)` syntax when fitting a hierarchical model with the`brm()` function. The terms before the `|` are those varying by group and you tell `brm()` what the grouping variable is after the `|`. In the case of multiple group-level parameters--which is the case with this model (i.e., both intercept and the `x_z` slope)--, this syntax also estimates correlations among the group-level parameters. Kruschke's model doesn't appear to include such a correlation. Happily, we can use the `(||)` syntax instead, which omits correlations among the group-level parameters. If you're curious about the distinction, fit the model both ways and explore the differences in the `print()` output. For more on the topic, see the *Group-level terms* subsection of the `brmsformula` section of the [**brms** reference manual](https://CRAN.R-project.org/package=brms/brms.pdf) [@brms2022RM].

```{r fit17.4}
fit17.4 <- brm(
  data = my_data,
  family = student,
  y_z ~ 1 + x_z + (1 + x_z || Subj),
  prior = c(prior(normal(0, 10), class = Intercept),
            prior(normal(0, 10), class = b),
            prior(normal(0, 1), class = sigma),
            # The next line is new
            prior(normal(0, 1), class = sd),
            prior(exponential(one_over_twentynine), class = nu)),
  chains = 4, cores = 4,
  seed = 17,
  stanvars = stanvar(1/29, name = "one_over_twentynine"),
  file = "fits/fit17.04")
```

Did you catch that `prior(normal(0, 1), class = sd)` line in the code? That's the prior we used for our hierarchical variance parameters, $\sigma_0$ and $\sigma_1$. Just like with the scale parameter, $\sigma$, we used the zero-mean half-normal distribution. By default, **brms** sets their left boundary to zero, which keeps the HMC algorithm from exploring negative variance values.

Anyway, here's the model `summary()`.

```{r}
summary(fit17.4)
```

### The posterior distribution: Shrinkage and prediction

Keeping in the same spirit of @sec-Interpreting-the-posterior-distribution, we'll make the plots of Figure 17.5 in two ways. First, we'll use our `make_beta_0()` and  `make_beta_1()` functions to transform the model coefficients.

```{r}
#| warning: false

draws <- as_draws_df(fit17.4)

sd_x <- sd(my_data$X)
sd_y <- sd(my_data$Y)
m_x  <- mean(my_data$X)
m_y  <- mean(my_data$Y)

draws <- draws |> 
  mutate(b_0 = make_beta_0(zeta_0 = b_Intercept,
                           zeta_1 = b_x_z,
                           sd_x   = sd_x,
                           sd_y   = sd_y,
                           m_x    = m_x,
                           m_y    = m_y),
         b_1 = make_beta_1(zeta_1 = b_x_z,
                           sd_x   = sd_x,
                           sd_y   = sd_y)) |> 
  select(.draw, b_0, b_1)

head(draws)
```

Here's the top panel of Figure 17.4.
  
```{r}
#| fig-width: 4.25
#| fig-height: 4

# How many posterior lines would you like?
n_lines <- 250

my_data |> 
  mutate(Subj = factor(Subj, levels = 25:1)) |> 
  
  ggplot(aes(x = X, y = Y)) +
  geom_abline(data = draws |> slice(1:n_lines),
              aes(intercept = b_0, slope = b_1, group = .draw),
              alpha = 1/5, color = "grey50", linewidth = 1/4) +
  geom_point(aes(color = Subj),
             alpha = 1/2) +
  geom_line(aes(group = Subj, color = Subj),
            linewidth = 1/4) +
  scale_y_continuous(breaks = seq(from = 50, to = 250, by = 50)) +
  scale_color_manual(values = beyonce_palette(41, n = 25, type = "continuous"), breaks = NULL) +
  labs(subtitle = eval(substitute(paste("Data from all units with", n_lines, "credible population-level\nregression lines")))) +
  coord_cartesian(xlim = c(40, 95),
                  ylim = c(30, 270))
```

Recall how we can use `coef()` to extract the `Subj`-specific parameters. But we'll want posterior draws rather than summaries, which requires `summary = FALSE`. It'll take a bit of wrangling to get the output in a tidy format. Once we're there, the plot code will be fairly simple.

```{r}
#| fig-width: 7
#| fig-height: 6

# First collect and wrangle the draws for the Subj-level intercept and slopes
c <- rbind(coef(fit17.4, summary = FALSE)$Subj[, , "Intercept"],
           coef(fit17.4, summary = FALSE)$Subj[, , "x_z"]) |> 
  data.frame() |> 
  set_names(1:25) |> 
  mutate(draw = rep(1:4000, times = 2),
         param = rep(c("Intercept", "Slope"), each = 4000)) |> 
  pivot_longer(cols = `1`:`25`, names_to = "Subj") |> 
  pivot_wider(names_from = param, values_from = value) |> 
  # Now we're ready to un-standardize the standardized coefficients
  mutate(b_0 = make_beta_0(zeta_0 = Intercept,
                           zeta_1 = Slope,
                           sd_x   = sd_x,
                           sd_y   = sd_y,
                           m_x    = m_x,
                           m_y    = m_y),
         b_1 = make_beta_1(zeta_1 = Slope,
                           sd_x   = sd_x,
                           sd_y   = sd_y))

# How many lines would you like?
n_lines <- 250

# Plot:
my_data |> 
  mutate(Subj = factor(Subj, levels = 25:1)) |> 
  
  ggplot(aes(x = X, y = Y)) +
  geom_abline(data = c |> filter(draw <= n_lines),
              aes(intercept = b_0, slope = b_1), 
              color = "grey50", linewidth = 1/4, alpha = 1/5) +
  geom_point(aes(color = Subj)) +
  scale_x_continuous(breaks = seq(from = 50, to = 90,  by = 20)) +
  scale_y_continuous(breaks = seq(from = 50, to = 250, by = 100)) +
  scale_color_manual(values = beyonce_palette(41, n = 25, type = "continuous"), breaks = NULL) +
  labs(subtitle = "Each unit now has its own bundle of credible regression lines") +
  coord_cartesian(xlim = c(45, 90),
                  ylim = c(50, 270)) +
  facet_wrap(~ Subj |> factor(levels = 1:25))
```

There's some good pedagogy in that method. But I like having options and in this case `fitted()` affords a simpler workflow. Here's the preparatory data wrangling step.

```{r}
# How many posterior lines would you like?
n_lines <- 250

# Since we're working with straight lines, we only need two x-values
nd <- tibble(x_z = c(-5, 5)) |> 
  # Transform the standardized `x_z` values to the unstandardized metric
  mutate(X    = x_z * sd(my_data$X) + mean(my_data$X),
         name = str_c("V", 1:n()))

f <- fitted(fit17.4,
            newdata = nd,
            # Since we only want the fixed effects, we'll use `re_formula` 
            # to marginalize over the random effects
            re_formula = Y_z ~ 1 + X_z,
            summary = FALSE,
            # Here we use `ndraws` to subset right from the get go
            ndraws = n_lines) |> 
  as_tibble() |> 
  mutate(draw = 1:n()) |> 
  pivot_longer(cols = -draw) |> 
  # Transform the standardized `y_z` values back into the unstandardized `Y` metric
  mutate(Y = value * sd(my_data$Y) + mean(my_data$Y)) |> 
  # Now attach the predictor values to the output
  left_join(nd, by = "name")

head(f)  
```

For the second time, here's the top panel of Figure 17.4, this time based off of `fitted()`.

```{r}
#| fig-width: 4.25
#| fig-height: 4

p1 <- my_data |> 
  mutate(Subj = factor(Subj, levels = 25:1)) |>
  
  ggplot(aes(x = X, y = Y)) +
  geom_line(data = f,
            aes(group = draw),
            alpha = 1/5, color = "grey50", linewidth = 1/4) +
  geom_point(aes(color = Subj),
             alpha = 1/2) +
  geom_line(aes(group = Subj, color = Subj),
            linewidth = 1/4) +
  scale_y_continuous(breaks = seq(from = 50, to = 250, by = 50)) +
  scale_color_manual(values = beyonce_palette(41, n = 25, type = "continuous")) +
  labs(subtitle = eval(substitute(paste("Data from all units with", n_lines, "credible population-level\nregression lines")))) +
  coord_cartesian(xlim = c(40, 95),
                  ylim = c(30, 270)) +
  theme(legend.position = "none")

p1
```

The whole process is quite similar for the `Subj`-specific lines. There are two main differences. First, we need to specify which `Subj` values we'd like to get `fitted()` points for. That goes into our `nd` tibble. Second, we omit the `re_formula` argument. There are other subtleties, like with the contents of the `bind_cols()` function. But hopefully those are self-evident.

```{r}
# How many posterior lines would you like?
n_lines <- 250

nd <- tibble(x_z = c(-5, 5)) |> 
  mutate(X = x_z * sd(my_data$X) + mean(my_data$X)) |> 
  expand_grid(Subj = distinct(my_data, Subj) |> pull()) |> 
  mutate(name = str_c("V", 1:n()))

f <- fitted(fit17.4,
            newdata = nd,
            summary = FALSE,
            ndraws = n_lines) |> 
  as_tibble() |> 
  mutate(draw = 1:n()) |> 
  pivot_longer(cols = -draw) |> 
  mutate(Y = value * sd(my_data$Y) + mean(my_data$Y)) |> 
  left_join(nd, by = "name")

head(f)  
```

And now for the second time, here's the bottom panel of Figure 17.4, this time based off of `fitted()`.

```{r}
#| fig-width: 7
#| fig-height: 10

p2 <- my_data |> 
  mutate(Subj = factor(Subj, levels = 25:1)) |>
  
  ggplot(aes(x = X, y = Y)) +
  geom_line(data = f,
            aes(group = draw),
            alpha = 1/5, color = "grey50", linewidth = 1/4) +
  geom_point(aes(color = Subj)) +
  scale_x_continuous(breaks = seq(from = 50, to = 90,  by = 20)) +
  scale_y_continuous(breaks = seq(from = 50, to = 250, by = 100)) +
  scale_color_manual(values = beyonce_palette(41, n = 25, type = "continuous"), breaks = NULL) +
  labs(subtitle = "Each unit now has its own bundle of credible regression lines") +
  coord_cartesian(xlim = c(45, 90),
                  ylim = c(50, 270)) +
  facet_wrap(~ Subj |> factor(levels = 1:25))

# Combine with patchwork
p3 <- plot_spacer()

p4 <- (p3 | p1 | p3) + 
  plot_layout(widths = c(1, 4, 1))

(p4 / p2) + plot_layout(heights = c(0.6, 1))
```

Especially if you're new to these kinds of models, it's easy to get lost in all that code. And for real--the wrangling required for those plots was no joke. The primary difficulty was that we had to convert standardized solutions to unstandardized solutions, which leads to an important distinction. When we used the first method of working with the `as_draws_df()` and `coef()` output, we focused on **transforming the model parameters**. In contrast, when we used the second method of working with the `fitted()` output, we focused instead on **transforming the model predictions and predictor values**. This distinction can be really confusing, at first. Stick with it! There will be times one method is more convenient or intuitive than the other. It's good to have both methods in your repertoire.

## Quadratic trend and weighted data

Quadratic models follow the general form

$$y = \beta_0 + \beta_1 x + \beta_2 x^2,$$

where $\beta_2$ is the quadratic term which, when 0, reduces the results to a simple linear model. That's right; the linear model is a special case of the quadratic.

This time the data come from the American Community Survey and Puerto Rico Community Survey. In his footnote #3, Kruschke indicated "Data are from [http://www.census.gov/hhes/www/income/data/Fam_Inc_SizeofFam1.xls](http://www.census.gov/hhes/www/income/data/Fam_Inc_SizeofFam1.xls), retrieved December 11, 2013. Median family income for years 2009-2011." As to our `read.csv()` code, note the `comment.char` argument.

```{r}
#| message: false

my_data <- read.csv("data.R/IncomeFamszState3yr.csv", comment.char = "#")

glimpse(my_data)
```

Here we'll standardize all variables but `State`, our grouping variable. It'd be silly to try to standardize that.

```{r}
my_data <- my_data |> 
  mutate(family_size_z   = standardize(FamilySize),
         median_income_z = standardize(MedianIncome),
         se_z            = SampErr / (mean(SampErr)))

glimpse(my_data)
```

With **brms**, there are a couple ways to handle measurement error on a variable (e.g., see [Chapter 14](https://bookdown.org/content/3890/missing-data-and-other-opportunities.html) of my ebook, [*Statistical rethinking with brms, ggplot2, and the tidyverse*](https://bookdown.org/content/3890/) [@kurzStatisticalRethinkingBrms2023]. Here we'll use the `se()` syntax, following the form `response | se(se_response, sigma = TRUE)`. In this form, `se` stands for standard error, the loose frequentist analogue to the Bayesian posterior $\textit{SD}$. Unless you're fitting a meta-analysis on summary information, make sure to specify `sigma = TRUE`. Without that you'll have no estimate for $\sigma$! For more information on the `se()` method, go to the [**brms** reference manual](https://CRAN.R-project.org/package=brms/brms.pdf) and find the *Additional response information* subsection of the `brmsformula` section [@brms2022RM, p. 42].

```{r fit17.5}
fit17.5 <- brm(
  data = my_data,
  family = student,
  median_income_z | se(se_z, sigma = TRUE) ~ 1 + family_size_z +  I(family_size_z^2) + 
    (1 + family_size_z +  I(family_size_z^2) || State),
  prior = c(prior(normal(0, 10), class = Intercept),
            prior(normal(0, 10), class = b),
            prior(normal(0, 1), class = sigma),
            prior(normal(0, 1), class = sd),
            prior(exponential(one_over_twentynine), class = nu)),
  chains = 4, cores = 4,
  stanvars = stanvar(1/29, name = "one_over_twentynine"),
  seed = 17,
  file = "fits/fit17.05")
```

Did you notice the `I(family_size_z^2)` part of the `formula`? The **brms** package follows a typical convention in **R** statistical functions in that if you want to multiply a variable by itself as in a quadratic model, you nest the `family_size_z^2` part within the `I()` function.

Take a look at the model summary.

```{r}
print(fit17.5)
```

Do see that `Ifamily_size_zE2` row? That's the summary of our quadratic term.

### Results and interpretation

A new model type requires a different approach to un-standardizing our standardized coefficients. Based on Equation 17.3, we can convert our coefficients using the formulas

\begin{align*}
\beta_0 & = \zeta_0 \operatorname{SD}_y + M_y - \frac{\zeta_1 M_x \operatorname{SD}_y}{\operatorname{SD}_x} + \frac{\zeta_2 M^{2}_x \operatorname{SD}_y}{\operatorname{SD}^{2}_x}, \\
\beta_1 & = \frac{\zeta_1 \operatorname{SD}_y}{\operatorname{SD}_x} - \frac{2 \zeta_2 M_x \operatorname{SD}_y}{\operatorname{SD}^{2}_x}, \text{and} \\
\beta_2 & = \frac{\zeta_2 \operatorname{SD}_y}{\operatorname{SD}^{2}_x}.
\end{align*}

We'll make new custom functions to use them.

```{r}
make_beta_0 <- function(zeta_0, zeta_1, zeta_2, sd_x, sd_y, m_x, m_y) {
  zeta_0 * sd_y + m_y - zeta_1 * m_x * sd_y / sd_x + zeta_2 * m_x^2 * sd_y / sd_x^2
}

make_beta_1 <- function(zeta_1, zeta_2, sd_x, sd_y, m_x) {
  zeta_1 * sd_y / sd_x - 2 * zeta_2 * m_x * sd_y / sd_x^2
}

make_beta_2 <- function(zeta_2, sd_x, sd_y) {
  zeta_2 * sd_y / sd_x^2
}

# Save the sample statistics
m_x  <- mean(my_data$FamilySize)
m_y  <- mean(my_data$MedianIncome)
sd_x <- sd(my_data$FamilySize)
sd_y <- sd(my_data$MedianIncome)
```

Now we'll extract our posterior draws and make the conversions.

```{r}
#| warning: false
#| fig-width: 7
#| fig-height: 6

draws <- as_draws_df(fit17.5) |> 
  mutate(b_0 = make_beta_0(zeta_0 = b_Intercept,
                           zeta_1 = b_family_size_z,
                           zeta_2 = b_Ifamily_size_zE2,
                           sd_x   = sd_x,
                           sd_y   = sd_y,
                           m_x    = m_x,
                           m_y    = m_y),
         b_1 = make_beta_1(zeta_1 = b_family_size_z,
                           zeta_2 = b_Ifamily_size_zE2,
                           sd_x   = sd_x,
                           sd_y   = sd_y,
                           m_x    = m_x),
         b_2 = make_beta_2(zeta_2 = b_Ifamily_size_zE2,
                           sd_x   = sd_x,
                           sd_y   = sd_y)) |> 
  select(.draw, b_0:b_2)
```

Our `geom_abline()` approach from before won't work with curves. We'll have to resort to `geom_line()`. With the `geom_line()` approach, we'll need many specific values of model-implied `MedianIncome` across a densely-packed range of `FamilySize`. We want to use a lot of `FamilySize` values, like 30 or 50 or so, to make sure the curves look smooth. Below, we'll use 50 (i.e., `length.out = 50`). But if it's still not clear why, try plugging in a lesser value, like 5 or so. You'll see.

```{r}
# How many posterior lines would you like?
n_lines <- 200

set.seed(17)
draws <- draws |> 
  slice_sample(n = n_lines) |> 
  rownames_to_column(var = "draw") |> 
  expand_grid(FamilySize = seq(from = 1, to = 9, length.out = 50)) |> 
  mutate(MedianIncome = b_0 + b_1 * FamilySize + b_2 * FamilySize^2)

head(draws)
```

Now we're ready to make the top panel of Figure 17.7.
                   
```{r}
#| fig-width: 4.25
#| fig-height: 4

my_data |>
  ggplot(aes(x = FamilySize, y = MedianIncome)) +
  geom_line(data = draws,
            aes(group = .draw),
            alpha = 1/5, color = "grey67", linewidth = 1/4) +
  geom_line(aes(group = State, color = State),
            alpha = 2/3, linewidth = 1/4) +
  geom_point(aes(color = State),
             alpha = 2/3, size = 1/2) +
  scale_x_continuous("Family size", breaks = 1:8) +
  scale_color_manual(values = beyonce_palette(41, n = 52, type = "continuous"), breaks = NULL) +
  labs(y = "Median income",
       title = "All states") +
  coord_cartesian(xlim = c(1, 8),
                  ylim = c(0, 150000))
```

Like before, we'll extract the group-level coefficients (i.e., those specific to the `State`s) with the `coef()` function. And also like before, the `coef()` output will require a little wrangling.

```{r}
# First collect and wrangle the draws for the State-level intercept and slopes
c <- rbind(coef(fit17.5, summary = FALSE)$State[, , "Intercept"],
           coef(fit17.5, summary = FALSE)$State[, , "family_size_z"],
           coef(fit17.5, summary = FALSE)$State[, , "Ifamily_size_zE2"]) |> 
  data.frame() |> 
  mutate(draw  = rep(1:4000, times = 3),
         param = rep(c("Intercept", "family_size_z", "Ifamily_size_zE2"), each = 4000)) |> 
  pivot_longer(cols = Alabama:Wyoming, names_to = "State") |> 
  pivot_wider(names_from = param, values_from = value)  |> 
  # Let's go ahead and make the standardized-to-unstandardized conversions, here
  mutate(b_0 = make_beta_0(zeta_0 = Intercept,
                           zeta_1 = family_size_z,
                           zeta_2 = Ifamily_size_zE2,
                           sd_x   = sd_x,
                           sd_y   = sd_y,
                           m_x    = m_x,
                           m_y    = m_y),
         b_1 = make_beta_1(zeta_1 = family_size_z,
                           zeta_2 = Ifamily_size_zE2,
                           sd_x   = sd_x,
                           sd_y   = sd_y,
                           m_x    = m_x),
         b_2 = make_beta_2(zeta_2 = Ifamily_size_zE2,
                           sd_x   = sd_x,
                           sd_y   = sd_y)) |> 
  # We just want the first 25 states, from Alabama through Mississippi, so we'll `filter()`
  filter(State <= "Mississippi")

str(c)
```

Now we'll subset by `n_lines`, `expand_grid()` by `FamilySize`, and use the model formula to compute the expected `MedianIncome` values.

```{r}
# How many posterior lines would you like?
n_lines <- 200

set.seed(17)
c <- c |> 
  group_by(State) |>
  slice_sample(n = n_lines) |>
  ungroup() |> 
  select(draw, State, b_0, b_1, b_2) |> 
  expand_grid(FamilySize = seq(from = 1, to = 9, length.out = 50)) |> 
  mutate(MedianIncome = b_0 + b_1 * FamilySize + b_2 * FamilySize^2)

head(c)
```

Finally, we're ready for the `State`-specific miniatures in Figure 17.7.

```{r}
#| fig-width: 7
#| fig-height: 6

my_data |>
  filter(State <= "Mississippi") |> 
 
  ggplot(aes(x = FamilySize, y = MedianIncome)) +
  geom_line(data = c,
            aes(group = draw),
            alpha = 1/5, color = "grey67", linewidth = 1/4) +
  geom_point(aes(color = State)) +
  geom_line(aes(color = State)) +
  scale_x_continuous("Family size", breaks = 1:8) +
  scale_color_manual(values = beyonce_palette(41, n = 52, type = "continuous"), breaks = NULL) +
  labs(y = "Median income",
       subtitle = "Each State now has its own bundle of credible regression curves.") +
  coord_cartesian(xlim = c(1, 8),
                  ylim = c(0, 150000)) +
  theme(legend.position = "none") +
  facet_wrap(~ State)
```

Magic! As our model coefficients proliferate, the `fitted()` approach from above starts to look more and more appetizing. Check it out for yourself.

Although "almost all of the posterior distribution [was] below $\nu = 4$" in the text (p. 500), the bulk of our $\nu$ distribution spanned across much larger values.

```{r}
#| fig-width: 3.5
#| fig-height: 2.75

as_draws_df(fit17.5) |>
  ggplot(aes(x = nu)) +
  stat_histinterval(point_interval = mode_hdi, .width = 0.95,
                    breaks = 40, color = bp[1], fill = bp[6],
                    normalize = "panels", slab_color = bp[5]) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(x = NULL,
       title = expression(Our~big~nu))
```

I'm guessing the distinction in our $\nu$ distribution and that in the text is our use of the `se()` syntax in the `brm()` `formula`. If you have a better explanation, [share it](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/issues).

### Further extensions

Kruschke discussed the ease with which users of Bayesian software might specify nonlinear models. Check out Bürkner's [-@Bürkner2022Non_linear] vignette, [*Estimating non-linear models with brms*](https://CRAN.R-project.org/package=brms/vignettes/brms_nonlinear.html), for more on the topic. Though I haven't used it, I believe it is also possible to use the $t$ distribution to model group-level variation in **brms** (see [this GitHub discussion](https://github.com/paul-buerkner/brms/issues/231) for details).

## Procedure and perils for expanding a model

Across several chapters, we've already dipped our toes into posterior predictive checks. For more on the PPC "double dipping" issue, check out Gelman's [*Discussion with Sander Greenland on posterior predictive checks*](http://andrewgelman.com/2014/08/11/discussion-sander-greenland-posterior-predictive-checks/) or Simpson's [*Touch me, I want to feel your data*](http://andrewgelman.com/2017/09/07/touch-want-feel-data/), which is itself connected to @gabry2019visualization, [*Visualization in Bayesian workflow*](https://arxiv.org/abs/1709.01449).

## 17.6. ~~Exercises~~ Bonus: Standardized coefficients can be effect sizes

Kruschke didn't put it this way in the text, but standardized regression coefficients can be seen as effect sizes. Specifically, the $\zeta_1$ coefficient in a continuous regression model with a single metric predictor is in a correlation metric, and Pearson's correlation coefficients are one of the most popular effect sizes used in the social sciences. Kruschke focused on robust Student $t$ regression in this chapter, but here we'll first work within a Gaussian framework, and then generalize to the Student $t$.

Back in @sec-Robust-linear-regression, we used the `height_weight_generator()` to make some standardized height and weight data. Here we repeat that code for the example. As before, we save the data as `d`.

```{r}
d <- height_weight_generator(n = 300, male_prob = 0.50, seed = 17) |> 
  mutate(height_z = standardize(height),
         weight_z = standardize(weight))

head(d)
```

Now fit a Gaussian version of `fit17.1`, which we'll call `fit17.1b`.

```{r fit17.1b}
fit17.1b <- brm(
  data = d,
  family = gaussian,
  weight_z ~ 1 + height_z,
  prior = c(prior(normal(0, 10), class = Intercept),
            prior(normal(0, 10), class = b),
            prior(normal(0, 1), class = sigma)),
  chains = 4, cores = 4,
  seed = 17,
  file = "fits/fit17.01b")
```

If you look at the `print()` output, you'll see it's very similar to the output for `fit17.1`, but with the omission of the summary for `nu` (which the Gaussian likelihood fixes to $\infty$).

```{r}
#| eval: false

# Execute this on your own computer
print(fit17.1)
print(fit17.1b)
```

Here we focus on the summary for $\zeta_1$.

```{r}
fixef(fit17.1b)["height_z", ] |> round(digits = 3)
```

Compare that to the Pearson's correlation coefficient from the `cor()` function.

```{r}
d |> 
  summarize(r = cor(weight_z, height_z))
```

Provided you use flat priors, they will the the same within MCMC sampling error. That is, $\zeta_1$ in a simple univariate Gaussian model fit to completely standardized data is in a correlation metric. Here's what that correlation effect size looks like in a plot.

```{r}
#| fig-width: 4.5
#| fig-height: 4

# How many posterior lines would you like?
n_lines <- 100

as_draws_df(fit17.1b) |> 
  slice(1:n_lines) |> 
  
  ggplot() +
  geom_abline(aes(intercept = b_Intercept, slope = b_height_z, group = .draw),
              alpha = 1/3, color = bp[2], linewidth = 1/4) +
  geom_point(data = d,
             aes(x = height_z, y = weight_z),
             alpha = 1/2, color = bp[5]) +
  scale_x_continuous("height (standardized)", expand = expansion(mult = 0.5)) +
  scale_y_continuous("weight (standardized)", expand = expansion(mult = 0.5)) +
  labs(subtitle = expression("Behold the effect size, "*italic(r)%~~%zeta[1]==0.45)) 
```

```{r}
#| fig-width: 4.5
#| fig-height: 4
#| echo: false
#| eval: false

# Figure 17.4 (for reference)

make_beta_0 <- function(zeta_0, zeta_1, sd_x, sd_y, m_x, m_y) {
  zeta_0 * sd_y + m_y - zeta_1 * m_x * sd_y / sd_x
}

make_beta_1 <- function(zeta_1, sd_x, sd_y) {
  zeta_1 * sd_y / sd_x
}

sd_x <- sd(d$height)
sd_y <- sd(d$weight)
m_x  <- mean(d$height)
m_y  <- mean(d$weight)

draws <- as_draws_df(fit17.1) |> 
  mutate(b_0 = make_beta_0(zeta_0 = b_Intercept,
                           zeta_1 = b_height_z,
                           sd_x   = sd_x,
                           sd_y   = sd_y,
                           m_x    = m_x,
                           m_y    = m_y),
         b_1 = make_beta_1(zeta_1 = b_height_z,
                           sd_x   = sd_x,
                           sd_y   = sd_y))

# How many posterior lines would you like?
n_lines <- 100

d |> 
  ggplot(aes(x = height, y = weight)) +
  geom_abline(data = draws |> slice(1:n_lines),
              aes(intercept = b_0, slope = b_1, group = .draw),
              alpha = 1/3, color = bp[2], linewidth = 1/4) + 
  geom_point(alpha = 1/2, color = bp[5]) +
  labs(x = "height",
       y = "weight",
       subtitle = eval(substitute(paste("Data with", n_lines, "credible regression lines")))) +
  coord_cartesian(xlim = c(50, 80),
                  ylim = c(-50, 470))
```

### Robust correlations with Robust $t$ regression

The Gaussian framework above generalized to $\zeta_1$ parameters from robust $t$ models. For example, compare the $\zeta_1$ summary with the Pearson's correlation from above.

```{r}
fixef(fit17.1)["height_z", ] |> round(digits = 3)
```

The value is very similar. But it is slightly higher than the Pearson's correlation coefficient, and it's also slightly larger than the $\zeta_1$ summary from the Gaussian model. In general, as the posterior for $\nu$ goes to infinity, $\zeta_1$ from a $t$ model will converge on both the $\zeta_1$ posterior from a Gaussian model, and also on the Pearson's correlation coefficient. However, as $\nu$ gets very small (indicating notably fatter tails than the conventional Gaussian), the posterior for $\zeta_1$ will generally increase. Such a $\zeta_1$ will still be in a correlation-like metric, but not that's not strictly identical to Pearson's $r$.

### Limitations

Personally, I really like this interpretation of $\zeta_1$, and I expect many of my more substantively-oriented colleagues would like it too. We psychologists like a good correlation. This approach also works well for small to moderately-sized correlations. However, it will get a little odd for very strong correlations. This is because whereas Pearson's $r$ is bounded between -1 and 1, there's nothing in the structure of a standardized univariable Gaussian or $t$ model to restrict the posterior for $\zeta_1$ to that range. To give a sense, let's look at a quick data example.

```{r}
set.seed(17)

d_cor <- data.frame(x = rnorm(n = 100)) |> 
  mutate(y = rnorm(n = n(), mean = 0 + 0.95 * x, sd = 0.31)) |> 
  mutate(across(.cols = everything(), .fns = standardize))

# What?
glimpse(d_cor)
```

The `d_cor` data has two variables, `x` and `y`, which have both been standardized using the `standardize()` function. Let's fit another simple Gaussian model, just like `fit17.1b`, above.

```{r fit17.6}
fit17.6 <- brm(
  data = d_cor,
  family = gaussian,
  y ~ 1 + x,
  prior = c(prior(normal(0, 10), class = Intercept),
            prior(normal(0, 10), class = b),
            prior(normal(0, 1), class = sigma)),
  chains = 4, cores = 4,
  seed = 17,
  file = "fits/fit17.06")
```

Before we look at the posterior, here's the Pearson's $r$.

```{r}
d_cor |> 
  summarize(r = cor(x, y))
```

That's a very strong correlation. Now take a close look at the summary for $\zeta_1$.

```{r}
fixef(fit17.6)["x", ] |> round(digits = 3)
```

The posterior mean is centered right on the Pearson's $r$, but look at the 95% intervals. The upper limit has crossed 1, which is the upper boundary for a Pearson's $r$. Even so, this model is still a good description of the data, which we can see in plot.

```{r}
#| fig-width: 4.5
#| fig-height: 4

# How many posterior lines would you like?
n_lines <- 50

as_draws_df(fit17.6) |> 
  slice(1:n_lines) |> 
  
  ggplot() +
  geom_abline(aes(intercept = b_Intercept, slope = b_x, group = .draw),
              alpha = 1/3, color = bp[2], linewidth = 1/4) +
  geom_point(data = d_cor,
             aes(x = x, y = y),
             alpha = 1/2, color = bp[5]) +
  scale_x_continuous("x (standardized)", expand = expansion(mult = 0.5)) +
  scale_y_continuous("y (standardized)", expand = expansion(mult = 0.5))
```

### More correlation talk

For more on Bayesian correlations and standardized univariable models, we turn to blog posts. For a series of blog posts on [robust] Bayesian correlations within the context of multivariate distributions, see @Baath2013The, @BaezOrtega2018Bayesian, and @Kurz2019Bayesian. For my follow-up blog post comparing multivariate models with $\zeta_1$ from univariate models like those in this chapter, see @Kurz2019Options.

## Session info {-}

```{r}
sessionInfo()
```

## Comments {-}

