---
title: "Chapter 25. Tools in the Trunk"
author: "A Solomon Kurz"
date: "`r format(Sys.Date())`"
output:
  github_document
---

```{r, echo = F}
knitr::opts_chunk$set(fig.align = "center")
```

# Tools in the Trunk

"This chapter includes some important topics that apply to many different models throughout the book... The sections can be read independently of each other and at any time" (p. 721)

## Reporting a Bayesian analysis

> Bayesian data analyses are not yet standard procedure in many fields of research, and no conventional format for reporting them has been established. Therefore, the researcher who reports a Bayesian analysis must be sensitive to the background knowledge of his or her specific audience, and must frame the description accordingly. (p. 721)

At the time of this writing (early 2020), this is still the case. See [*Discussion points for Bayesian inference*](https://www.researchgate.net/publication/338849264_Discussion_points_for_Bayesian_inference) for a recent discussion from several Bayesian scholars.

### Essential points.

> Recall the basic steps of a Bayesian analysis from Section 2.3 (p. 25): Identify the data, define a descriptive model, specify a prior, compute the posterior distribution, interpret the posterior distribution, and, check that the model is a reasonable description of the data. Those steps are in logical order, with each step building on the previous step. That logical order should be preserved in the report of the analysis. (p. 722)

Kruschke then gave recommendations for motivating Bayesian inference. His (2017) paper with Liddell, [*The Bayesian New Statistics: Hypothesis testing, estimation, meta-analysis, and power analysis from a Bayesian perspective*](https://link.springer.com/article/10.3758/s13423-016-1221-4), might be helpful in this regard. Many of the other points Kruschke made in this section (e.g., adequately reporting the data structure, the priors, evidence for convergence) can be handled by adopting open science practices.

If your data and research questions are simple and straightforward, you might find it easy to detail these and other concerns in the primary manuscript. The harsh reality is many journals place tight constraints on word and/or page limits. If your projects are not of the simple and straightforward type, supplemental materials are your friend. Regardless of a journal's policy on hosting supplemental materials on the official journal website, you can detail your data, priors, MCMC diagnostics, and all the other fine-grained details of your analysis in supplemental documents hosted in publicly-accessible repositories like the [Open Science Framework (OSF)](https://osf.io/). If possible, do consider making your data openly available. Regardless of the status of your data, please consider making all your **R** scripts available as supplementary material. To reiterate from Chapter 3, I strongly recommend checking out [R Notebooks](https://bookdown.org/yihui/rmarkdown/notebook.html) for that purpose. They are a type of R Markdown document with augmentations that make them more useful for working scientists. You can learn more about them [here](https://rstudio-pubs-static.s3.amazonaws.com/256225_63ebef4029dd40ef8e3679f6cf200a5a.html) and [here](https://www.r-bloggers.com/why-i-love-r-notebooks-2/). And for a more comprehensive overview, check out Xie, Allaire, and Grolemund's [*R Markdown: The Definitive Guide*](https://bookdown.org/yihui/rmarkdown/).

### Optional points.

For more thoughts on robustness checks, check out a couple Gelman's blog posts, [*What's the point of a robustness check?*](https://statmodeling.stat.columbia.edu/2017/11/29/whats-point-robustness-check/) and [*Robustness checks are a joke*](https://statmodeling.stat.columbia.edu/2018/11/14/robustness-checks-joke/), along with the action in the comments section. 

In addition to posterior predictive checks, which are great (see [Kruschke, 2013](https://jkkweb.sitehost.iu.edu/articles/Kruschke2013BJMSP.pdf)), consider prior predictive checks, too. For a great introduction to the topic, check out Gabry, Simpson, Vehtari, Betancourt, and Gelman's [*Visualization in Bayesian workflow*](https://rss.onlinelibrary.wiley.com/doi/pdf/10.1111/rssa.12378).

### Helpful points.

For more ideas on open data, check out Rouder's [*The what, why, and how of born-open data*](https://link.springer.com/article/10.3758/s13428-015-0630-z). You might also check out Klein and colleagues' [*A practical guide for transparency in psychological science*](https://lirias.kuleuven.be/1999530?limo=0) and Martone, Garcia-Castro, and VandenBos's [*Data sharing in psychology*](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5920518/).

As to posting your model fits, this could be done in any number of ways, including as official supplemental materials hosted by the journal, on GitHub, or on the OSF. At a base level, this means saving your fits as external files. We've already been modeling this with our `brm()` code throughout this book. With the `save` argument, we saved the model fits within the [`fits` folder on GitHub](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/tree/master/fits). You might adopt a similar approach. But do be warned: **brms** fit objects contain a copy of the data used to create them. For example, here's how we might reload `fit24.1` from last chapter.

```{r, message = F}
fit24.1 <- readRDS("fits/fit24.01.rds")
```

By indexing the fit object with `$data`, you can see the data.

```{r, warning = F, message = F}
library(tidyverse)
library(brms)

fit24.1$data %>% 
  glimpse()
```

Here's a quick way to remove the data from the fit object.

```{r}
fit24.1$data <- NULL
```

Confirm it worked.

```{r}
fit24.1$data
```

Happily, the rest of the information is still there for you. E.g., here's the summary.

```{r}
print(fit24.1)
```

## Functions for computing highest density intervals

You can find a copy of Kruschke's scripts, including `DBDA2E-utilities.R`, at [https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/tree/master/data.R](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/tree/master/data.R).

### R code for computing HDI of a grid approximation.

> We can imagine the grid approximation of a distribution as a landscape of poles sticking up from each point on the parameter grid, with the height of each pole indicating the probability mass at that discrete point. We can imagine the highest density region by visualizing a rising tide: We gradually flood the landscape, monitoring the total mass of the poles that protrude above water, stopping the flood when 95% (say) of the mass remains protruding. The waterline at that moment defines the highest density region (e.g., [Hyndman, 1996](https://amstat.tandfonline.com/doi/abs/10.1080/00031305.1996.10474359#.XkGooxNKg1I)). (p. 725)

```{r}
HDIofGrid <- function(probMassVec, credMass = 0.95) {
  
  # Arguments:
  #   probMassVec is a vector of probability masses at each grid point.
  #   credMass is the desired mass of the HDI region.
  # Return value:
  #   A list with components:
  #   indices is a vector of indices that are in the HDI
  #   mass is the total mass of the included indices
    #   height is the smallest component probability mass in the HDI
  # Example of use: For determining HDI of a beta(30,12) distribution
  #   approximated on a grid:
  #   > probDensityVec = dbeta( seq(0,1,length=201) , 30 , 12 )
  #   > probMassVec = probDensityVec / sum( probDensityVec )
  #   > HDIinfo = HDIofGrid( probMassVec )
  #   > show( HDIinfo )
  
  sortedProbMass <- sort(probMassVec, decreasing = TRUE)
  HDIheightIdx <- min(which(cumsum(sortedProbMass) >= credMass))
  HDIheight <- sortedProbMass[HDIheightIdx]
  HDImass <- sum(probMassVec[probMassVec >= HDIheight])
  
  return(list(indices = which(probMassVec >= HDIheight),
              mass    = HDImass, height = HDIheight))
  
}
```

I found Kruschke's description of his `HDIofGrid()` a bit opaque. Happily, we can understand this function with a little help from an example posted at [https://rdrr.io/github/kyusque/DBDA2E-utilities/man/HDIofGrid.html](https://rdrr.io/github/kyusque/DBDA2E-utilities/man/HDIofGrid.html).

```{r}
prob_density_vec <- dbeta(seq(0, 1, length = 201), 30, 12)
prob_mass_vec    <- prob_density_vec / sum(prob_density_vec)
HDI_info         <- HDIofGrid(prob_mass_vec)

show(HDI_info)
```

To walk that through a bit, `prob_density_vec` is a vector of density values for $\operatorname{beta} (30, 12)$ based on 201 evenly-spaced values spanning the parameter space for $\theta$ (i.e., from 0 to 1). In the second line, we converted those density values to the probability metric by dividing each by their sum, which we then saved as `prob_mass_vec`. In the third line we shoved those probability values into Kruschke's `HDIofGrid()` and saved the results as `HDI_info`. The output of the fourth line, `show(HDI_info)`, showed us the results (i.e., the contents of `HDI_info`).

As to those results, the values in saved as `$indices` are the row numbers for all cases in `prob_mass_vec` that were within the HDI. The value in `$mass` showed the actual width of the HDI. Because we're only working with finite samples (i.e., `length = 201`), we won't likely get a perfect 95% HDI. The value in `$height` is the density value for *the waterline that defines the highest density region*. A plot might make that less abstract.

```{r, fig.width = 7, fig.height = 2.75}
# wrangle
tibble(row     = 1:length(prob_density_vec),
       theta   = seq(0, 1, length = length(prob_density_vec)),
       density = prob_mass_vec,
       cred    = if_else(row %in% HDI_info$indices, 1, 0)) %>% 
  
  # plot
  ggplot(aes(x = theta, y = density)) +
  # HDI
  geom_ribbon(data = . %>% 
                filter(cred == 1),
              aes(ymin = 0, ymax = density),
              fill = "grey50") +
  # density line
  geom_line(color = "grey33") +
  # waterline
  geom_hline(yintercept = HDI_info$height, linetype = 2, color = "skyblue") +
  # fluff
  annotate(geom = "text", x = .2, y = 0.0046,
           label = '"waterline" that defines all points\ninside the highest density region') +
  annotate(geom = "text", x = .715, y = 0.01,
           label = "95.28% HDI", color = "grey92", size = 5) +
  xlab(expression(theta)) +
  theme(panel.grid = element_blank())
```

### HDI of unimodal distribution is shortest interval.

> The algorithms [in the next sections] find the HDI by searching among candidate intervals of mass $M$. The shortest one found is declared to be the HDI. It is an approximation, of course. See [Chen and Shao (1999)](https://www.researchgate.net/publication/2442323_Monte_Carlo_Estimation_of_Bayesian_Credible_and_HPD_Intervals) for more details, and [Chen, He, Shao, and Xu (2003)](https://www.researchgate.net/publication/264969946_A_Monte_Carlo_gap_test_in_computing_HPD_regions) for dealing with the unusual situation of multimodal distributions. (p. 727)

### R code for computing HDI of a MCMC sample.

In this section, Kruschke provided the code for his `HDIofMCMC()` function. We recreate it, below, with a few mild formatting changes.

```{r}
HDIofMCMC <- function(sampleVec, credMass = .95) {
  
  # Computes highest density interval from a sample of representative values,
  #   estimated as shortest credible interval.
  # Arguments:
  #   sampleVec
  #     is a vector of representative values from a probability distribution.
  #   credMass
  #     is a scalar between 0 and 1, indicating the mass within the credible
  #     interval that is to be estimated.
  # Value:
  #   HDIlim is a vector containing the limits of the HDI
  
  sortedPts <- sort(sampleVec)
  ciIdxInc <- ceiling(credMass * length(sortedPts))
  nCIs <- length(sortedPts) - ciIdxInc
  ciWidth <- rep(0, nCIs)
  for (i in 1:nCIs) {
    ciWidth[i] <- sortedPts[i + ciIdxInc] - sortedPts[i]
  }
  HDImin <- sortedPts[which.min(ciWidth)]
  HDImax <- sortedPts[which.min(ciWidth) + ciIdxInc]
  HDIlim <- c(HDImin, HDImax)
  
  return(HDIlim)
  
}
```

Let's continue working with `fit24.1` to see how Kruschke's `HDIofMCMC()` works. First we need to extract the posterior draws.

```{r}
post <- posterior_samples(fit24.1)
```

Here's how you might use the function to get the HDIs for the first hierarchical variance parameter.

```{r}
HDIofMCMC(post$sd_Eye__Intercept)
```

Kruschke's `HDIofMCMC()` works very much the same as the summary functions from **tidybayes**. For example, here's good old `tidybayes::mode_hdi()`.

```{r, warning = F, message = F}
library(tidybayes)

mode_hdi(post$sd_Eye__Intercept)
```

If you'd like to use **tidybayes** to just pull the HDIs without the extra information, just use the `hdi()` function.

```{r}
hdi(post$sd_Eye__Intercept)
```

Just in case you're curious, Kruschke's `HDIofMCMC()` function returns the same information as `tidybayes::hdi()`. Let's confirm.

```{r}
HDIofMCMC(post$sd_Eye__Intercept) == hdi(post$sd_Eye__Intercept)
```

Identical.

### R code for computing HDI of a function.

> The function described in this section finds the HDI of a unimodal probability density function that is specified mathematically in R. For example, the function can find HDI's of normal densities or of beta densities or of gamma densities, because those densities are specified as functions in R. (p. 728).

If you recall, we've been using this funciton off and on since Chapter 4. Here is it, again, with mildly reformated code and parameter names.

```{r}
hdi_of_icdf <- function(name, width = .95, tol = 1e-8, ... ) {
  
  # Arguments:
  #   `name` is R's name for the inverse cumulative density function
  #   of the distribution.
  #   `width` is the desired mass of the HDI region.
  #   `tol` is passed to R's optimize function.
  # Return value:
  #   Highest density iterval (HDI) limits in a vector.
  # Example of use: For determining HDI of a beta(30, 12) distribution, type
  #   `hdi_of_icdf(qbeta, shape1 = 30, shape2 = 12)`
  #   Notice that the parameters of the `name` must be explicitly stated;
  #   e.g., `hdi_of_icdf(qbeta, 30, 12)` does not work.
  # Adapted and corrected from Greg Snow's TeachingDemos package.
  
  incredible_mass <-  1.0 - width
  interval_width <- function(low_tail_prob, name, width, ...) {
    name(width + low_tail_prob, ...) - name(low_tail_prob, ...)
  }
  opt_info <- optimize(interval_width, c(0, incredible_mass), 
                       name = name, width = width, 
                       tol = tol, ...)
  hdi_lower_tail_prob <- opt_info$minimum
  
  return(c(name(hdi_lower_tail_prob, ...),
           name(width + hdi_lower_tail_prob, ...)))
  
}
```

Here's how it works for the standard normal distribution.

```{r}
hdi_of_icdf(qnorm, mean = 0, sd = 1)
```

By default, it returns 95% HDIs. Here's how it'd work if you wanted the 80% intervals for $\operatorname{beta}(2, 2)$.

```{r}
hdi_of_icdf(qbeta, shape1 = 2, shape2 = 2, width = .8)
```

## Reparameterization

> There are situations in which one parameterization is intuitive to express a distribution, but a different parameterization is required for mathematical convenience. For example, we may think intuitively of the standard deviation of a normal distribution, but have to parameterize the distribution in terms of the precision (i.e., reciprocal of the variance). (p. 729)

The details in the rest of this section are beyond the scope of this project.

## Censored Data in ~~JAGS~~ brms

"In many situations some data are censored, which means that their values are known only within a certain range" (p. 732) Happily, **brms** is capable of handling censored variables. The setup is a little different from how Kruschke described for JAGS. From the [**brms** reference manual](https://cran.r-project.org/web/packages/brms/brms.pdf) (section `brmsformula`), we read:

> With the exception of categorical, ordinal, and mixture families, left, right, and interval censoring can be modeled through `y | cens(censored) ~ predictors`. The censoring variable (named `censored` in this example) should contain the values `'left'`, `'none'`, `'right'`, and `'interval'` (or equivalently `-1`, `0`, `1`, and `2`) to indicate that the corresponding observation is left censored, not censored, right censored, or interval censored. For interval censored data, a second variable (let's call it `y2`) has to be passed to `cens`. In this case, the formula has the structure `y | cens(censored,y2) ~ predictors`. While the lower bounds are given in `y`, the upper bounds are given in `y2` for interval censored data. Intervals are assumed to be open on the left and closed on the right: `(y,y2]`.

We'll make sense of all this in just a moment. First, let's see how Kruschke described the example in the text.

> To illustrate why it is important to include censored data in the analysis, consider a case in which $N = 500$ values are generated randomly from a normal distribution with $\mu = 100$ and $\sigma = 15$. Suppose that values above 106 are censored, as are values in the interval between 94 and 100. For the censored values, all we know is the interval in which they occurred, but not their exact value. (p. 732)

I'm now aware that we have access to Kruschke's censored data, so we'll just make our own based on his description. We'll start off by simulating the idealized uncensored data, `y`, based on $\operatorname{Normal} (100, 15)$.

```{r}
n <- 500

set.seed(25)

d <- tibble(y = rnorm(n, mean = 100, sd = 15))
```

To repeat, Kruschke described two kinds of censoring:

* "values above 106 are censored", 
* "as are values in the interval between 94 and 100."

This leaves us with three thresholds. For simplicity, we'll just name them `t1`, `t2` and `t3`, with their order based on their numeric values.

```{r}
t1 <-  94
t2 <- 100
t3 <- 106
```

In the last sentence in the block quote from the **brms** reference manual, we learned "intervals are assumed to be open on the left and closed on the right: `(y,y2]`." It's a little unclear, to me, if this is how Kruschke defined his intervals, but since we're working with **brms** we'll just use this convention. Thus, we will define "values in the interval between 94 and 100" as `y >= t1 & y < t2`. We will define "values above 106" as `y > t3`.

```{r}
d <-
  d %>% 
  mutate(y1  = if_else(y >= t1 & y < t2, t1,
                       if_else(y > t3, t3, y)),
         y2  = if_else(y >= t1 & y < t2, t2, y),
         cen = if_else(y >= t1 & y < t2, "interval",
                       if_else(y > t3, "right", "none")))

d
```

First look at the new `cen` column. When the values in `y` are not censored, we see `"none"`. Otherwise, `cen` indicates if they are right censored (i.e., `"right"`) or interval censored (i.e., `"interval"`). We used those exact terms based on the block quote from the **brms** reference manual. Now look at `y1`. When `cen == "interval"`, those values are the same as the original column `y`. The same goes for `y2`. Otherwise, the `y1` column contains the relevant lower thresholds values. That is, when `cen == "interval"`, we see the value for `t1` (i.e., 94). When `cen == "right"`, we see the value for `t3` (i.e., 106). For the interval-censored rows, the values in `y2` contain the values for the upper threshold (i.e., `t2`, which is 100). But when the rows are right censored, the values in `y2` are simply the same as the original `y` values. In the rows where `cen == "right"`, it really doesn't matter what values you put in the `y2` column as long as they aren't `NA`. This is because **brms** will only reference them for rows in which `cen == "interval"`.

I would not spend any time trying to equate this with Kruschke’s exposition at the top of page 734. This is a different coding method from what you might use for JAGS. Let’s make one more data change. Here we’ll make a new variable, `y_na`, that only has values for which `cen == "none"`.

```{r}
d <-
  d %>% 
  mutate(y_na  = ifelse(cen == "none", y, NA))

d
```

In the text, Kruschke reported he had 255 uncensored values (p. 732). Here's the breakdown of our data.

```{r}
d %>% 
  count(cen)
```

We got really close! Let's look at what we've done with a couple histograms.

```{r, fig.width = 6, fig.height = 2.5, warning = F}
d %>% 
  pivot_longer(c(y, y_na)) %>% 
  
  ggplot(aes(x = value)) +
  geom_histogram(size = .25, binwidth = 2.5, fill = "grey67", color = "grey92") +
  xlab(NULL) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~name, ncol = 2)
```

Here's how to fit the first model, which only uses the uncensored values.

```{r fit25.1}
# define the stanvars
mean_y <- mean(d$y_na, na.rm = T)
sd_y   <- sd(d$y_na, na.rm = T)

stanvars <- 
  stanvar(mean_y, name = "mean_y") + 
  stanvar(sd_y,   name = "sd_y")

# fit the model
fit25.1 <-
  brm(data = d,
      family = gaussian,
      y_na ~ 1,
      prior = c(prior(normal(mean_y, sd_y * 100), class = Intercept),
                prior(normal(0, sd_y), class = sigma)),
      chains = 4, cores = 4,
      stanvars = stanvars, 
      file = "fits/fit25.01.rds")
```

Check the summary for the naïve model.

```{r}
print(fit25.1)
```

Relative to the true data-generating process for the original variable `y`, those parameters look pretty biased. Now let's practice fitting censored model.

This model is one of the rare occasions where we'll set out initial values for the model intercept. In my first few attempts, `brm()` had great difficulty initializing the chains using the default initial values. We'll help it out by setting them at `mean_y`. Recall that when you set custom initial values in **brms**, you save them in a list with the number of lists equaling the number of HMC chains. Because we're using the default `chains = 4`, well need four lists of intercept start values, `mean_y`. You can set them to different values, if you'd like.

```{r fit25.2}
inits <- list(Intercept = mean_y)

inits_list <- list(inits, inits, inits, inits)

fit25.2 <-
  brm(data = d,
      family = gaussian,
      y1 | cens(cen, y2) ~ 1,
      prior = c(prior(normal(mean_y, sd_y * 100), class = Intercept),
                prior(normal(0, sd_y), class = sigma)),
      chains = 4, cores = 4,
      stanvars = stanvars, 
      inits = inits_list,  # here we insert our start values for the intercept
      file = "fits/fit25.02.rds")
```

Now check the summary for the model accounting for the censoring.

```{r}
print(fit25.2)
```

All 500 cases were used (`Number of observations: 500`) and the model did a great job capturing the data-generating parameters. Before we can make our version of Figure 25.4, we'll need to extract the posterior draws. We'll start with `fit25.1`.

```{r}
post <- 
  posterior_samples(fit25.1) %>% 
  mutate(mu = b_Intercept,
         `(mu-100)/sigma` = (b_Intercept - 100) / sigma)

head(post)
```

These subplots look a lot like those from back in Section 16.2. Since this is the last plot of the book, it seems like we should make the effort to stitch all the subplots together with **patchwork**. To reduce some of the code redundancy with the six subplots of the marginal posteriors, we'll make a custom geom, `geom_hist()`.

```{r}
geom_hist <- function(xintercept = xintercept, binwidth = binwidth, ...) {
  
  list(
    geom_vline(xintercept = xintercept, color = "white", size = 1),
    geom_histogram(fill = "grey67", color = "grey92", size = .2, binwidth = binwidth),
    stat_pointintervalh(aes(y = 0), point_interval = mode_hdi, .width = .95),
    scale_y_continuous(NULL, breaks = NULL),
    theme(panel.grid = element_blank())
  )
  
}
```

Now we have our `geom_hist()`, here are the first three histograms for the marginal posteriors from `fit25.1`.

```{r}
p1 <-
  post %>% 
  ggplot(aes(x = mu)) +
  geom_hist(xintercept = 100, binwidth = 0.25) +
  xlab(expression(mu))

p3 <-
  post %>% 
  ggplot(aes(x = sigma)) +
  geom_hist(xintercept = 15, binwidth = 0.1) +
  xlab(expression(sigma))

p4 <-
  post %>% 
  ggplot(aes(x = `(mu-100)/sigma`)) +
  geom_hist(xintercept = 0, binwidth = 0.025) +
  xlab(expression((mu-100)/sigma))
```

The histogram of the censored data with the posterior predictive density curves superimposed will take a little more work.

```{r, fig.width = 3, fig.height = 2}
n_lines <- 50

p2 <-
  post %>% 
  mutate(iter = 1:n()) %>% 
  slice(1:n_lines) %>% 
  expand(nesting(mu, sigma, iter),
         y_na = seq(from = 40, to = 120, by = 1)) %>% 
  mutate(density = dnorm(x = y_na, mean = mu, sd = sigma)) %>% 
  
  ggplot(aes(x = y_na)) + 
  geom_histogram(data = d, 
                 aes(y = stat(density)),
                 color = "grey92", fill = "grey67",
                 size = .2, binwidth = 2.5) +
  geom_line(aes(y = density, group = iter),
            size  = 1/4, alpha = 1/3, color = "grey25") +
  scale_x_continuous("data with posterior predictive lines", limits = c(40, 110)) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())
```

Now extract the posterior draws from our censored model, `fit25.2`, and repeat the process.

```{r}
post <- 
  posterior_samples(fit25.2) %>% 
  mutate(mu = b_Intercept,
         `(mu-100)/sigma` = (b_Intercept - 100) / sigma)

p5 <-
  post %>% 
  ggplot(aes(x = mu)) +
  geom_hist(xintercept = 100, binwidth = 0.15) +
  xlab(expression(mu))

p7 <-
  post %>% 
  ggplot(aes(x = sigma)) +
  geom_hist(xintercept = 15, binwidth = 0.1) +
  xlab(expression(sigma))

p8 <-
  post %>% 
  ggplot(aes(x = `(mu-100)/sigma`)) +
  geom_hist(xintercept = 0, binwidth = 0.01) +
  xlab(expression((mu-100)/sigma))

p6 <-
  post %>% 
  mutate(iter = 1:n()) %>% 
  slice(1:n_lines) %>% 
  expand(nesting(mu, sigma, iter),
         y_na = seq(from = 40, to = 120, by = 1)) %>% 
  mutate(density = dnorm(x = y_na, mean = mu, sd = sigma)) %>% 
  
  ggplot(aes(x = y_na)) + 
  geom_histogram(data = d, 
                 aes(y = stat(density)),
                 color = "grey92", fill = "grey67",
                 size = .2, binwidth = 2.5) +
  geom_line(aes(y = density, group = iter),
            size  = 1/4, alpha = 1/3, color = "grey25") +
  scale_x_continuous("data with posterior predictive lines", limits = c(40, 110)) +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank())
```

Load **patchwork**, combine the subplots, and annotate a bit.

```{r, fig.width = 6, fig.height = 7.25, warning = F}
library(patchwork)

((p1 | p2) / (p3 | p4) / (p5 | p6) / (p7 | p8)) +
  plot_annotation(title = "This is our final plot, together.",
                  caption = expression(atop(italic("Upper quartet")*": Censored data omitted from analysis; parameter estimates are too small.  ", italic("Lower quartet")*": Censored data imputed in known bins; parameter estimates are accurate."))) &
  theme(plot.caption = element_text(hjust = 0))
```

## What Next?

"If you have made it this far and you are looking for more, you might peruse posts at [Kruschke's] blog, [[https://doingbayesiandataanalysis.blogspot.com/]](https://doingbayesiandataanalysis.blogspot.com/), and search there for topics that interest you." In addition to the other references Kruschke mentioned, you might also check out McElreath's [*Statistical Rethinking*](https://xcelab.net/rm/statistical-rethinking/). The first edition came out in 2015 and the second is scheduled to be released sometime in early/mid 2020. Much like this project, I have recoded *Statistical Rethinking* in a **bookdown** form, [here](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/). You can also find other tutorial material at my academic blog, [https://solomonkurz.netlify.com/post/](https://solomonkurz.netlify.com/post/).

## Reference {-}

[Kruschke, J. K. (2015). *Doing Bayesian data analysis, Second Edition: A tutorial with R, JAGS, and Stan.* Burlington, MA: Academic Press/Elsevier.](https://sites.google.com/site/doingbayesiandataanalysis/)

## Session info {-}

```{r}
sessionInfo()
```

```{r, message = F, warning = F, echo = F}
# remove our objects
rm(fit24.1, HDIofGrid, prob_density_vec, prob_mass_vec, HDI_info, HDIofMCMC, post, hdi_of_icdf, n, d, t1, t2, t3, mean_y, sd_y, stanvars, fit25.1, inits, inits_list, fit25.2, geom_hist, p1, p3, p4, n_lines, p2, p5, p7, p8, p6)
```

```{r, echo = F, message = F, warning = F, results = "hide"}
pacman::p_unload(pacman::p_loaded(), character.only = TRUE)

ggplot2::theme_set(ggplot2::theme_grey())

bayesplot::color_scheme_set("blue")
```

