# What and why {.unnumbered}

Kruschke began his text with "This book explains how to actually do Bayesian data analysis, by real people (like you), for realistic data (like yours)." In the same way, this project is designed to help those real people do Bayesian data analysis. My contribution is converting Kruschke's JAGS and Stan code for use in Bürkner's [**brms** package](https://github.com/paul-buerkner/brms) [@R-brms; @burknerBrmsPackageBayesian2017; @burknerAdvancedBayesianMultilevel2018], which makes it easier to fit Bayesian regression models in **R** [@R-base] using Hamiltonian Monte Carlo. I also prefer plotting and data wrangling with the packages from the [**tidyverse**](http://style.tidyverse.org) [@R-tidyverse; @wickhamWelcomeTidyverse2019]. So we'll be using those methods, too.

This ebook is not meant to stand alone. It's a supplement to the second edition of Kruschke's [-@kruschkeDoingBayesianData2015] [*Doing Bayesian data analysis: A tutorial with R, JAGS, and Stan*](https://sites.google.com/site/doingbayesiandataanalysis/). Please give the source material some love.

## **R** setup

To get the full benefit from this ebook, you'll need some software. Happily, everything will be free (provided you have access to a decent personal computer and an good internet connection).

First, you'll need to install **R**, which you can learn about at [https://cran.r-project.org/](https://cran.r-project.org/).

Though not necessary, your **R** experience might be more enjoyable if done through the free RStudio interface, which you can learn about at [https://rstudio.com/products/rstudio/](https://rstudio.com/products/rstudio/).

Once you have installed **R**, execute the following to install the bulk of the add-on packages. This will probably take a few minutes to finish. Go make yourself a coffee.

```{r}
#| eval: false

packages <- c("bayesplot", "brms", "coda", "cowplot", "cubelyr", "devtools", "fishualize", "GGally", "ggdist", "ggExtra", "ggforce", "ggmcmc", "ggridges", "ggthemes", "janitor", "lisa", "loo", "palettetown", "patchwork", "psych", "remotes", "rstan", "santoku", "scico", "tidybayes", "tidyverse")

install.packages(packages, dependencies = TRUE)
```

A few of the other packages are not officially available via the Comprehensive R Archive Network (CRAN; https://cran.r-project.org/). You can download them directly from GitHub by executing the following.

```{r}
#| eval: false

remotes::install_github("clauswilke/colorblindr")
devtools::install_github("dill/beyonce")
devtools::install_github("ropenscilabs/ochRe")
```

It's possible you'll have problems installing some of these packages. Here are some likely suspects and where you can find help:

* for difficulties installing **brms**, go to [https://github.com/paul-buerkner/brms#how-do-i-install-brms](https://github.com/paul-buerkner/brms#how-do-i-install-brms) or search around in the [**brms** section of the Stan forums](https://discourse.mc-stan.org/c/interfaces/brms/36); and
* for difficulties installing **rstan**, go to [https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started).

## We have updates

For a brief rundown of the version history, we have:

### Version 0.1.0

I released the 0.1.0 version of this project in February 17, 2020. It was the first [fairly] complete draft including material from all the chapters in Kruschke's text. The supermajority of Kruschke's JAGS and Stan models were fit **brms** 2.11.5. The results were saved in the [`fits` folder on GitHub](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/tree/master/fits) and most of the results are quite comparable to those in the original text. We also reproduced most of the data-related figures and tables and little subpoints and examples sprinkled throughout Kruschke's prose.

### Version 0.2.0

The 0.2.0 update came in May 19, 2020. Noteworthy changes included:

* reproducing the simulation necessary for Figure 7.3 (see [GitHub issue #14](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/issues/14)) with help from Cardy Moten III ([\@cmoten](https://github.com/cmoten));
* with guidance from Bjørn Peare Bartholdy ([\@bbartholdy](https://github.com/bbartholdy)), Mladen Jovanović ([\@mladenjovanovic](https://github.com/mladenjovanovic)), Cory Whitney ([\@CWWhitney](https://github.com/CWWhitney)), and Brenton M. Wiernik ([\@bwiernik](https://github.com/bwiernik)), we improved in-text citations and reference sections using [BibTex](http://www.bibtex.org/) [@BibTeX2020], [Better BibTeX](https://github.com/retorquere/zotero-better-bibtex) [@BetterBibTeXZotero2020], and [zotero](https://www.zotero.org/) [@ZoteroYourPersonal2020];
* the plot resolution increased with `fig.retina = 2.5`; and
* small code, hyperlink, and typo corrections.

### Version 0.3.0

The 0.3.0 update came in September 22, 2020. Noteworthy changes included:

* adding the [Kruschke-style model diagrams](https://solomonkurz.netlify.app/post/make-model-diagrams-kruschke-style/) throughout the text (e.g., Figure 8.5 in @sec-Example-Difference-of-biases);
* adding chapter-specific plotting schemes with help from the [**cowplot** package](https://wilkelab.org/cowplot) [@R-cowplot], Wilke's [-@wilkeFundamentalsDataVisualization2019] [*Fundamentals of data visualization*](https://clauswilke.com/dataviz/), and many other great color-scheme packages; 
* an overhaul to the plotting workflow in @sec-Prior-knowledge-expressed-as-a-beta-distribution; and
* updating all model fits with **brms** version 2.13.5.

### Version 0.4.0

The 0.4.0 update came in May 6, 2021. Noteworthy changes included:

* using the Metropolis algorithm to fit the bivariate Bernoulli model for Figure 7.6 (@sec-The-posterior-via-the-Metropolis-algorithm), thanks to help from [Omid Ghasemi](https://github.com/OmidGhasemi21);
* corrections to mistakes around the `lag()` and `lead()` functions in @sec-MCMC-accuracy;
* an added bonus section clarifying the pooled standard deviation for standardized mean differences (@sec-Bonus-Pooled-standard-deviation);
* refining the custom `stat_wilke()` plotting function in @sec-Metric-Predicted-Variable-with-Multiple-Metric-Predictors;
* an overhaul of the bonus section covering effect sizes;
* refining/correcting the threshold workflow for univariable logistic regression models (@sec-Dichotomous-Predicted-Variable);
* fixing the divergent transitions issue for the robust logistic regression model by adding boundaries on the prior (@sec-Robust-logistic-regression);
* corrections to a few incorrectly computed effect sizes in @sec-Ordinal-Predicted-Variable;
* the addition of a new bonus section (@sec-Bonus-Consider-the-interceps-only-softmax-model) highlighting the benefits of the intercepts-only softmax model;
* expansions to the material on censored data (@sec-Censored-Data-in-brms) and the addition of a brief introduction to truncated data (@sec-Bonus-Truncation); and
* updating all HMC fits to the current version of **brms** (2.15.0).

### Version 1.0.0

The big 1.0.0 update came in May 4, 2022. This was the first full draft in the sense that it contained **brms** versions of all of Kruschke's JAGS and Stan models, excluding examples not currently possible with the **brms** paradigm (e.g., @sec-Well-cover-information-criteria-instead). Noteworthy changes included:

* two new solutions for the conditional logistic models of @sec-Nominal-Predicted-Variable thanks to the generous efforts by [Henrik Singmann](https://github.com/singmann) and [Mattan Ben-Shachar](https://github.com/mattansb/);
* replacing the depreciated `posterior_samples()` function with the new `posterior::as_draws_df()`-based workflow;
* adding a new solution for the multivariate Bernoulli model with different trial numbers via the `resp_subset()` function in @sec-Uneven-multivariate-Benoulli-via-the-resp_subset-approach;
* improving the efficiency of the intercept-only Bernoulli models with the new `lb` and `ub` arguments for priors of `class = Intercept` (e.g., `fit7.1a`, `fit7.1b`, and `fit7.2` in @sec-Markov-Chain-Monte-Carlo);
* updating all model fits with **brms** version 2.17.0; and
* various minor code, hyperlink, and typo corrections.

### Version 1.1.0

The 1.1.0 update came in January 25, 2023. Noteworthy changes included:

* replacing my incorrect use of `tidyr::expand()` with a more appropriate `tidyr::expand_grid()` workflow, thanks to insights from [Desislava Petkova](https://github.com/dipetkov);
* adopting the new `linewidth` argument for several **ggplot2** geoms (see [here](https://www.tidyverse.org/blog/2022/08/ggplot2-3-4-0-size-to-linewidth/)); and
* minor prose and code edits throughout.

### Version 1.2.0

The 1.2.0 update came in August 13, 2025. Noteworthy changes included:

* switching to a [Quarto ebook](https://quarto.org/docs/books/) format;
* updating the code to better reflect [tidyverse-style guidelines](https://style.tidyverse.org/);
* switching to the base pipe (`|>`);
* correcting minor mistakes in the density examples of @sec-Discrete-distributions-Probability-mass;
* updating the grid-based workflow for the primary figures in @sec-Bayes-Rule;
* updating the analytic workflow for the primary figures in @sec-Inferring-a-Binomial-Probability-via-Exact-Mathematical-Analysis;
* clarifying and correcting some of the panels in Figure 10.3 (see @sec-Solution-by-grid-approximation) thanks to help from John Kruschke ([\@kruschke](https://github.com/kruschke));
* solving the HMC proposal schematics presented in @sec-HMC-sampling, thanks to help from John Kruschke;
* cleaning up some of the custom function code; and
* minor prose and code edits throughout.

### Version 1.3.0

Welcome to version 1.3.0! Noteworthy changes include:

* an expansion of and challenge to the effect size/Cohen's $d$ material in @sec-Metric-Predicted-Variable-on-One-or-Two-Groups and @sec-Metric-Predicted-Variable-with-One-Nominal-Predictor;
* including the computations from @sec-Bayes-rule-intuited;
* a small update to the geom default oriented code in @sec-Linear-function-of-a-single-metric-predictor; and
* a few other minor prose and code edits throughout.

### Some minor issues remain

There are some minor improvements I'd like to add in future versions. Most importantly, a few simulations, figures, and models are beyond my current skill set. I've opened separate GitHub issues for the most important ones and they are as follows:

* the effective-sample-size simulations in Section 7.5.2 and the corresponding plots in Figures 7.13 and 7.14 ([issue #15](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/issues/15)),
* several of the simulations in Sections 11.1.4, 11.3.1, and 11.3.2 and their corresponding figures (issues [#16](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/issues/16), [#17](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/issues/17), [#18](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/issues/18), and [#19](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/issues/19)),
* the stopping-rule simulations in Section 13.3.2 and their corresponding figures ([issue #20](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/issues/20)).

If you know how to conquer any of these unresolved challenges, I'd love to hear all about it. In addition, please feel free to open a new [GitHub issue](https://github.com/ASKurz/Doing-Bayesian-Data-Analysis-in-brms-and-the-tidyverse/issues) if you find any flaws in the other sections of the ebook.

## Thank-you's are in order

Before we enter the primary text, I'd like to thank the following for their helpful contributions:

* Bjørn Peare Bartholdy ([\@bbartholdy](https://github.com/bbartholdy)),
* David Baumeister ([\@xdavebx](https://github.com/xdavebx)),
* Mattan Ben-Shachar ([\@mattansb](https://github.com/mattansb/)),
* Paul-Christian Bürkner ([\@paul-buerkner](https://github.com/paul-buerkner)),
* Christopher Flach ([\@flachboard](https://github.com/flachboard)),
* Andrew Gelman ([\@andrewgelman](https://github.com/andrewgelman)),
* Omid Ghasemi ([\@OmidGhasemi21](https://github.com/OmidGhasemi21)),
* Mladen Jovanović ([\@mladenjovanovic](https://github.com/mladenjovanovic)),
* Matthew Kay ([\@mjskay](https://github.com/mjskay)),
* John K. Kruschke ([\@kruschke](https://github.com/kruschke)),
* TJ Mahr ([\@tjmahr](https://github.com/tjmahr)),
* Cardy Moten III ([\@cmoten](https://github.com/cmoten)),
* Lukas Neugebauer ([\@LukasNeugebauer](https://github.com/LukasNeugebauer)),
* Demetri Pananos ([\@Dpananos](https://github.com/dpananos)),
* Desislava Petkova ([\@dipetkov](https://github.com/dipetkov)),
* Peter Ralph ([\@petrelharp](https://github.com/petrelharp)),
* Henrik Singmann ([\@singmann](https://github.com/singmann)),
* Jacob B. Socolar ([\@jsocolar](https://github.com/jsocolar)),
* Aki Vehtari ([\@avehtari](https://github.com/avehtari)),
* Matti Vuorre ([\@mvuorre](https://github.com/mvuorre)),
* Cory Whitney ([\@CWWhitney](https://github.com/CWWhitney)), and
* Brenton M. Wiernik ([\@bwiernik](https://github.com/bwiernik)).

## License and citation

This book is licensed under the Creative Commons Zero v1.0 Universal license. You can learn the details, [here](https://github.com/ASKurz/Applied-Longitudinal-Data-Analysis-with-brms-and-the-tidyverse/blob/master/LICENSE). In short, you can use my work. Just please give me the appropriate credit the same way you would for any other scholarly resource. Here's the citation information:

```default
@book{kurzDoingBayesianDataAnalysis2025,
  title = {Doing {{Bayesian}} data analysis in brms and the tidyverse},
  author = {Kurz, A. Solomon},
  year = {2025},
  month = {12},
  edition = {Version 1.3.0},
  url = {https://solomon.quarto.pub/doing-bayesian-data-analysis-in-brms-and-the-tidyverse/}
}
```

